---
title: "Dynamic analysis"
output:
  html_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par(mgp = c(2, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin3 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", mai = rep(.1, 4))
    else NULL
  }
)

knitr::opts_chunk$set(echo       = TRUE,
                      margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center")

l <- "en_US.UTF-8"
Sys.setenv(LANGAGE = l)
Sys.setlocale(locale = l)
Sys.setlocale("LC_MESSAGES", l)
```

## Preamble

In this document, we perform the dynamic risk assessment. In particular, we investigate how the transient CRS dynamics and 30-year CRS burden change when we allow modest improvements in routine RCV coverage over time and introduce supplemental immunization activities (SIAs). 

Relevant simulations must be performed in `model_simulators.Rmd` prior to running this R Markdown document. 

## Installation of packages

Here are the packages required: 

```{r include=FALSE}
required <- c("tidyverse", "sf", "magrittr", "cowplot", "RColorBrewer", "lemon")
to_install <- setdiff(required, row.names(installed.packages()))
if (length(to_install)) install.packages(to_install)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(magrittr)
library(cowplot)
library(RColorBrewer)
library(lemon)
```

## Links to local data sets

Here we use the following links to local data:

```{r}
DATA_FOLDER <- file.path("..", "data")
PARAMETERS_FOLDER <- file.path("parameters")
POPULATION_DATA_FOLDER <- file.path(DATA_FOLDER, "population_data")
SPATIAL_POLYGONS_FOLDER <- file.path(DATA_FOLDER, "spatial_polygons")
VACCINATION_DATA_FOLDER <- file.path(DATA_FOLDER, "vaccination_data")

# derived data from `nigeria_data_organization_exploration.Rmd`
ng1_path <- file.path(SPATIAL_POLYGONS_FOLDER, "ng1.rds")
ng2_pop_path <- file.path(POPULATION_DATA_FOLDER, "ng2_pop.rds")

# local data
ng1_fertility_data_path <- file.path(POPULATION_DATA_FOLDER, "ng1_fertility_data.csv")
ng1_zone_mapping_path <- file.path(POPULATION_DATA_FOLDER, "ng1_zone_designations.xlsx")
ng1_mcv1_data_path <- file.path(VACCINATION_DATA_FOLDER, "ng1_mcv1_data.xlsx")
```

These are the paths we use for the simulation output. 

```{r}
ALL_OUTPUT_FOLDER <- "~/Desktop/output/"
STATIC_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "static_risk_simulations", "raw_simulations")
DYNAMIC_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "dynamic_risk_simulations", "raw_simulations")
SUMMARY_DYNAMIC_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "dynamic_risk_simulations", "summary_simulations")
SUMMARY_STATIC_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "static_risk_simulations", "summary_simulations")
FIGURES_FOLDER <- file.path("figures", "dynamic_analysis")
```

Change these paths if you would like to run this on your local device.

## Data sets

Let's read in the main data sets:

```{r}
ng2_pop <- readRDS(ng2_pop_path)
ng1_pop <- ng2_pop %>% 
  group_by(state, gender, agegrp) %>% 
  summarise(n=sum(n, na.rm=TRUE), .groups="keep") %>% 
  spread(key="gender", value="n") %>% 
  mutate(n=m+f) %>% 
  ungroup()
ng1_zone_mapping <- readxl::read_excel(ng1_zone_mapping_path)
ng1 <- readRDS(ng1_path) %>%
  left_join(ng1_zone_mapping, by="state")
ng1_mcv1_data <- readxl::read_excel(ng1_mcv1_data_path)
ng1_cbr_data <- read.csv(ng1_fertility_data_path) %>% 
  gather(key="agegrp", value="asfr", 2:ncol(.)) %>% 
  mutate(agegrp=substr(gsub("X", "", agegrp), 1, 2)) %>%
  right_join(ng1_pop, by=c("state", "agegrp")) %>% 
  mutate(asfr=ifelse(is.na(asfr), 0, asfr), 
         births=asfr*f) %>% 
  group_by(state) %>% 
  summarise(cbr=sum(births)/sum(n)*1000, .groups="keep") %>% 
  as.data.frame()
```

## Preliminary simulation output data management

We start by organizing the simulation output. 

### CRS burden: simple dynamic analysis (improvements in routine coverage)

We start by computing the CRS incidence and birth time series for each state. 

```{r eval=FALSE}
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(0, 0.01, 0.025, 0.05, 0.1)
all_basic_scenarios_crs_incidence_list <- list()
for (this_state in states) {
  print(this_state)
  crs_incidence_list <- list()
  for (ii in seq_along(scenarios)) {
    # retrieve the simulation output for the desired state
    if (scenarios[ii]==0) {
      output_files <- list.files(path=STATIC_SIMS_FOLDER, pattern=paste(this_state, "tf30", "vac", sep="_"))
      num_files <- length(output_files)
      for (jj in 1:min(num_files, 100)) {
        output_files[[jj]] <- paste(this_state, "tf30", "vac", paste0("simNum", jj), sep="_")
      }
    } else {
      output_files <- list.files(path=DYNAMIC_SIMS_FOLDER,
                                 pattern=paste(this_state, "tf30", paste0("vacRate", scenarios[ii]*1000, "_"), sep="_"))
      num_files <- length(output_files)
      for (jj in 1:num_files) {
        output_files[[jj]] <- paste(this_state, "tf30", paste0("vacRate", scenarios[ii]*1000), paste0("simNum", jj), sep="_")
      }
    }
    
    # retrieve the crs incidence and births in each simulation
    params <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "basic_parameters.rds"))
    state_crs_incidence_list <- list()
    pb = txtProgressBar(min = 0, max = length(output_files), initial = 0, style=3)
    for (jj in seq_along(output_files)) {
      if (scenarios[ii]==0) sim <- readRDS(file.path(STATIC_SIMS_FOLDER, output_files[[jj]]))
      else sim <- readRDS(file.path(DYNAMIC_SIMS_FOLDER, output_files[[jj]]))
      
      # compute crs incidence
      EP <- sim[, paste("EP", 1:params$num_stages, sep="")] %>% as.data.frame()
      sigma <- params$sigma
      c <- 0.65
      crs_incidence <- apply(EP, 1, function(x) sum(c*x*sigma))
      state_crs_incidence_list[[jj]] <- crs_incidence
      setTxtProgressBar(pb, jj)
    }
    close(pb)
    crs_incidence_list[[ii]] <- do.call(cbind, state_crs_incidence_list) %>% 
      as.data.frame() %>% 
      mutate(time=sim[, 1]) %>% 
      select(time, dplyr::everything()) %>% 
      setNames(c("time", paste("sim", seq_along(output_files), sep="")))
  }
  all_basic_scenarios_crs_incidence_list[[this_state]] <- crs_incidence_list
}
saveRDS(all_basic_scenarios_crs_incidence_list, file=file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "all_basic_scenarios_crs_incidence_list.rds"))
```

Let's compute the 30-year CRS burden for each simulation in each state under each scenario.

```{r eval=FALSE}
compute_crs_burden_distribution <- function(this_state, crs_incidence_list, births_list, scenarios) {
  state_crs_incidence_list <- crs_incidence_list[[this_state]]
  state_births <- births_list[[this_state]]
  
  # compute the long-term CRS burden in each simulation
  output_list <- list()
  for (jj in seq_along(scenarios)) {
    state_crs_incidence <- state_crs_incidence_list[[jj]]
    crs_burdens <- rep(NA, ncol(state_crs_incidence)-1)
    for (ii in 1:(ncol(state_crs_incidence)-1)) {
      crs_incidence <- state_crs_incidence[, 1+ii]
      births <- state_births[, 2]
      crs_burdens[ii] <- sum(crs_incidence)/sum(births)*10^5
    }
    output_list[[jj]] <- data.frame(scenario=scenarios[jj], state=this_state, crs_burden=crs_burdens, sim=seq_along(crs_burdens))
  }
  output <- do.call(rbind, output_list)
  return(output)
}
births_list <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "births_series_list.rds"))
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(0, 0.01, 0.025, 0.05, 0.1) 
crs_burden_dist <- do.call(rbind, purrr::map(states, function(x) 
  compute_crs_burden_distribution(x, crs_incidence_list=all_basic_scenarios_crs_incidence_list, births_list=births_list, scenarios=scenarios)))
saveRDS(crs_burden_dist, file=file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "simple_dynamic_crs_burden_dist.rds"))
```

Let's also read in the simulations performed for each state at the median $R_0$. 

```{r eval=FALSE}
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(-1, 0, 0.01, 0.025, 0.05, 0.1)
median_r0_simNums <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "median_r0_simNums_optimization.rds"))
ng0_median_r0_simNum <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "ng0_r0_simNums_optimization.rds"))
median_r0_simNums <- c(median_r0_simNums, "Nigeria"=as.numeric(ng0_median_r0_simNum["median"]))
basic_scenarios_medR0_crs_incidence_list <- list()
for (this_state in states) {
  print(this_state)
  median_r0_simNum <- median_r0_simNums[this_state]
  crs_incidence_list <- list()
  for (ii in seq_along(scenarios)) {
    # retrieve the simulation output for the desired state
    if (scenarios[ii]==-1) {
      output_file <- file.path(STATIC_SIMS_FOLDER, paste0(this_state, "_tf30", "_noVac", "_simNum", median_r0_simNum))
    } else {
      output_file <- file.path(DYNAMIC_SIMS_FOLDER, paste0(this_state, "_tf30", "_vacRate", scenarios[ii]*1000, "_simNum", median_r0_simNum))
    }
    
    # retrieve the crs incidence and births in each simulation
    params <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "basic_parameters.rds"))
    sim <- readRDS(output_file)
    
    # compute crs incidence
    EP <- sim[, paste("EP", 1:params$num_stages, sep="")] %>% as.data.frame()
    sigma <- params$sigma
    c <- 0.65
    crs_incidence <- apply(EP, 1, function(x) sum(c*x*sigma))
    crs_incidence_list[[ii]] <- crs_incidence %>% 
      as.data.frame() %>% 
      mutate(time=sim[, 1]) %>% 
      select(time, dplyr::everything()) %>% 
      setNames(c("time", paste0("s", ii)))
  }
  scenario_names <- c("baseline", 0, 0.01, 0.025, 0.05, 0.1)
  names(scenario_names) <- paste("s", 1:6, sep="")
  basic_scenarios_medR0_crs_incidence_list[[this_state]] <- purrr::reduce(crs_incidence_list, left_join, by="time") %>% 
    gather(key="scenario", value="crs", 2:ncol(.)) %>% 
    mutate(scenario=scenario_names[scenario], 
           state=this_state, simNum=median_r0_simNum) %>% 
    select(state, simNum, time, scenario, crs)
}
saveRDS(basic_scenarios_medR0_crs_incidence_list, file=file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "basic_scenarios_medR0_crs_incidence_list.rds"))
```

### CRS burden: advanced dynamic analysis (SIAs)

We start by computing the CRS incidence and birth time series for each state. 

```{r eval=FALSE}
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(0, 0.01, 0.025, 0.05, 0.1)
sia_types <- c("sia1", "sia2")
for (this_sia in sia_types) {
  all_advanced_scenarios_crs_incidence_list <- list()
  for (this_state in states) {
    print(this_state)
    crs_incidence_list <- list()
    for (ii in seq_along(scenarios)) {
      # retrieve the simulation output for the desired state
      output_files <- list.files(path=DYNAMIC_SIMS_FOLDER,
                                 pattern=paste(this_state, "tf30", this_sia, paste0("vacRate", scenarios[ii]*1000, "_"), sep="_"))
      num_files <- length(output_files)
      for (jj in 1:min(100, num_files)) {
        output_files[[jj]] <- paste(this_state, "tf30", this_sia, paste0("vacRate", scenarios[ii]*1000), paste0("simNum", jj), sep="_")
      }
      
      # retrieve the crs incidence and births in each simulation
      params <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "basic_parameters.rds"))
      state_crs_incidence_list <- list()
      pb = txtProgressBar(min = 0, max = length(output_files), initial = 0, style=3)
      for (jj in seq_along(output_files)) {
        sim <- readRDS(file.path(DYNAMIC_SIMS_FOLDER, output_files[[jj]]))
        
        # compute crs incidence
        EP <- sim[, paste("EP", 1:params$num_stages, sep="")] %>% as.data.frame()
        sigma <- params$sigma
        c <- 0.65
        crs_incidence <- apply(EP, 1, function(x) sum(c*x*sigma))
        state_crs_incidence_list[[jj]] <- crs_incidence
        setTxtProgressBar(pb, jj)
      }
      close(pb)
      crs_incidence_list[[ii]] <- do.call(cbind, state_crs_incidence_list) %>% 
        as.data.frame() %>% 
        mutate(time=sim[, 1]) %>% 
        select(time, dplyr::everything()) %>% 
        setNames(c("time", paste("sim", seq_along(output_files), sep="")))
    }
    all_advanced_scenarios_crs_incidence_list[[this_state]] <- crs_incidence_list
  }
  saveRDS(all_advanced_scenarios_crs_incidence_list, file=file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, paste0("all_advanced_scenarios_", this_sia, "_crs_incidence_list.rds")))
}
```

Let's compute the 30-year CRS burden for each simulation in each state under each scenario.

```{r eval=FALSE}
compute_crs_burden_distribution <- function(this_state, crs_incidence_list, births_list, scenarios) {
  state_crs_incidence_list <- crs_incidence_list[[this_state]]
  state_births <- births_list[[this_state]]
  
  # compute the long-term CRS burden in each simulation
  output_list <- list()
  for (jj in seq_along(scenarios)) {
    state_crs_incidence <- state_crs_incidence_list[[jj]]
    crs_burdens <- rep(NA, ncol(state_crs_incidence)-1)
    for (ii in 1:(ncol(state_crs_incidence)-1)) {
      crs_incidence <- state_crs_incidence[, 1+ii]
      births <- state_births[, 2]
      crs_burdens[ii] <- sum(crs_incidence)/sum(births)*10^5
    }
    output_list[[jj]] <- data.frame(scenario=scenarios[jj], state=this_state, crs_burden=crs_burdens, sim=seq_along(crs_burdens))
  }
  output <- do.call(rbind, output_list)
  return(output)
}
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(0, 0.01, 0.025, 0.05, 0.1) 
for (this_sia in c("sia1", "sia2")) {
  all_advanced_scenarios_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, paste0("all_advanced_scenarios_", this_sia, "_crs_incidence_list.rds")))
  advanced_dynamic_crs_burden_dist <- do.call(rbind, purrr::map(states, function(x) 
    compute_crs_burden_distribution(x, 
                                    crs_incidence_list=all_advanced_scenarios_crs_incidence_list, 
                                    births_list=births_list, 
                                    scenarios=scenarios))) %>% 
    as.data.frame() %>% 
    mutate(sia=this_sia)
  saveRDS(advanced_dynamic_crs_burden_dist, file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, paste0("advanced_dynamic_", this_sia, "_crs_burden_dist.rds")))
}
```

Let's also read in the simulations performed for each state at the median $R_0$. 

```{r eval=FALSE}
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(-1, 0, 0.01, 0.025, 0.05, 0.1)
median_r0_simNums <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "median_r0_simNums_optimization.rds"))
ng0_median_r0_simNum <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "ng0_r0_simNums_optimization.rds"))
median_r0_simNums <- c(median_r0_simNums, "Nigeria"=as.numeric(ng0_median_r0_simNum["median"]))
for (this_sia in c("sia1", "sia2")) {
  advanced_scenarios_medR0_crs_incidence_list <- list()
  for (this_state in states) {
    print(this_state)
    median_r0_simNum <- median_r0_simNums[this_state]
    crs_incidence_list <- list()
    for (ii in seq_along(scenarios)) {
      # retrieve the simulation output for the desired state
      if (scenarios[ii]==-1) {
        output_file <- file.path(STATIC_SIMS_FOLDER, paste0(this_state, "_tf30", "_noVac", "_simNum", median_r0_simNum))
      } else {
        output_file <- file.path(DYNAMIC_SIMS_FOLDER, paste0(this_state, "_tf30_", this_sia, "_vacRate", scenarios[ii]*1000, "_simNum", median_r0_simNum))
      }
      
      # retrieve the crs incidence and births in each simulation
      params <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "basic_parameters.rds"))
      sim <- readRDS(output_file)
      
      # compute crs incidence
      EP <- sim[, paste("EP", 1:params$num_stages, sep="")] %>% as.data.frame()
      sigma <- params$sigma
      c <- 0.65
      crs_incidence <- apply(EP, 1, function(x) sum(c*x*sigma))
      crs_incidence_list[[ii]] <- crs_incidence %>% 
        as.data.frame() %>% 
        mutate(time=sim[, 1]) %>% 
        select(time, dplyr::everything()) %>% 
        setNames(c("time", paste0("s", ii)))
    }
    scenario_names <- c("baseline", 0, 0.01, 0.025, 0.05, 0.1)
    names(scenario_names) <- paste("s", 1:6, sep="")
    advanced_scenarios_medR0_crs_incidence_list[[this_state]] <- purrr::reduce(crs_incidence_list, left_join, by="time") %>% 
      gather(key="scenario", value="crs", 2:ncol(.)) %>% 
      mutate(scenario=scenario_names[scenario], 
             state=this_state, simNum=median_r0_simNum, 
             sia=this_sia) %>% 
      select(sia, state, simNum, time, scenario, crs)
  }
  saveRDS(advanced_scenarios_medR0_crs_incidence_list, file=file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, 
                                                                   paste0("advanced_scenarios_", this_sia, "_medR0_crs_incidence_list.rds")))
}
```

## Simulation datasets

Here, we read in the organized simulation output data. 

### CRS burden: baseline analysis (without SIAs and improvement over time)

```{r}
baseline_crs_incidence_list <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "crs_incidence_series_list.rds"))
baseline_crs_burden_dist <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "crs_burden_dist.rds"))
births_list <- readRDS(file.path(SUMMARY_STATIC_SIMS_FOLDER, "births_series_list.rds"))
```

### CRS burden: simple dynamic analysis (improvements in routine coverage)

```{r}
all_basic_scenarios_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "all_basic_scenarios_crs_incidence_list.rds"))
simple_dynamic_crs_burden_dist <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "simple_dynamic_crs_burden_dist.rds"))
basic_scenarios_medR0_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "basic_scenarios_medR0_crs_incidence_list.rds"))
```

### CRS burden: advanced dynamic analysis (SIAs)

```{r}
all_advanced_scenarios_sia1_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "all_advanced_scenarios_sia1_crs_incidence_list.rds"))
advanced_dynamic_sia1_crs_burden_dist <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "advanced_dynamic_sia1_crs_burden_dist.rds"))
advanced_scenarios_sia1_medR0_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "advanced_scenarios_sia1_medR0_crs_incidence_list.rds"))
all_advanced_scenarios_sia2_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "all_advanced_scenarios_sia2_crs_incidence_list.rds"))
advanced_dynamic_sia2_crs_burden_dist <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "advanced_dynamic_sia2_crs_burden_dist.rds"))
advanced_scenarios_sia2_medR0_crs_incidence_list <- readRDS(file.path(SUMMARY_DYNAMIC_SIMS_FOLDER, "advanced_scenarios_sia2_medR0_crs_incidence_list.rds"))
```

## CRS burden: simple dynamic analysis (improvements in routine coverage)

In this section, we focus on simple dynamic analysis looking at how modest increases in routine rubella vaccination coverage over time diminishes the risk of RCV introduction even at coverage levels that are expected to increase CRS burden in the long run. 

### Yearly CRS incidence

Here, we look at the mean time series of CRS incidence in each state to get an idea of the possible temporal dynamics.

```{r}
# It takes the state name and returns a time series of the yearly CRS burden with uncertainty bars. 
plot_yearly_crs_dynamics <- function(this_state, crs_incidence_list, births_list, scenarios, 
                                     baseline_crs_incidence_list, baseline_births_list) {
  scenarios <- c(-1, scenarios)
  state_crs_incidence_list <- crs_incidence_list[[this_state]]
  state_births <- births_list[[this_state]]
  baseline_state_crs_incidence <- baseline_crs_incidence_list[[this_state]]
  baseline_state_births <- baseline_births_list[[this_state]]
  state_crs_incidence_list <- c(list(baseline_state_crs_incidence), state_crs_incidence_list)
  
  # compute the yearly crs incidence (per 100,000 live births) in each simulation
  yearly_incidence_summary_list <- list()
  for (this_scenario in seq_along(scenarios)) {
    state_crs_incidence <- state_crs_incidence_list[[this_scenario]]
    
    yearly_incidence_list <- list()
    yearly_incidence <- rep(0, floor(nrow(state_crs_incidence)/365))
    for (ii in 1:(ncol(state_crs_incidence)-1)) {
      crs_incidence <- state_crs_incidence[, 1+ii]
      if (this_scenario==1) births <- baseline_state_births[, 2]
      else births <- state_births[, 2]
      for (jj in seq(1, length(crs_incidence)-1, 365)) {
        yearly_incidence[floor(jj/365)+1] <- sum(crs_incidence[jj:(jj+364)])/sum(births[jj:(jj+364)])*10^5
      }
      yearly_incidence_list[[ii]] <- yearly_incidence
    }
    state_yearly_incidence <- do.call(cbind, yearly_incidence_list)
    state_yearly_incidence_summary <- apply(state_yearly_incidence, 1, function(x) 
      c(median(x), mean(x), sd(x), quantile(x, c(0.025, 0.975)))) %>% 
      t() %>% 
      as.data.frame() %>% 
      mutate(scenario=this_scenario, year=seq(1, floor(nrow(state_crs_incidence)/365), 1)) %>% 
      select(scenario, year, dplyr::everything()) %>% 
      setNames(c("scenario", "year", "median", "mean", "sd", "l95", "u95"))
    yearly_incidence_summary_list[[this_scenario]] <- state_yearly_incidence_summary
  }
  yearly_incidence_summary <- do.call(rbind, yearly_incidence_summary_list) %>% 
    mutate(scenario=factor(scenario, (-1):6))
  
  p <- ggplot(yearly_incidence_summary) + 
    geom_line(mapping=aes(x=year, y=mean, color=scenario), size=1) + 
    geom_point(mapping=aes(x=year, y=mean, color=scenario), size=1) +
    labs(x="Year", y="CRS cases/\n100,000 live births", title=this_state) + 
    scale_x_continuous(breaks=seq(0, 40, 5)) +
    scale_color_manual(values=brewer.pal(length(scenarios), "Dark2"), labels=c("No vaccination", paste0(scenarios[-1]*100, "%"))) + 
    theme_bw() + 
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          panel.background=element_blank(), panel.border=element_blank(),
          axis.line = element_line(colour = "black")) + 
    theme(axis.text = element_text(size=12), axis.title = element_text(size=14))
  return(p)
}
```

```{r echo=FALSE}
states <- c(sort(unique(pull(ng2_pop, state))), "Nigeria")
scenarios <- c(0, 0.01, 0.025, 0.05, 0.1) 
plots <- purrr::map(states, function(x) 
  plot_yearly_crs_dynamics(this_state=x, crs_incidence_list=all_basic_scenarios_crs_incidence_list, births_list=births_list, scenarios=scenarios, 
                           baseline_crs_incidence_list=baseline_crs_incidence_list, baseline_births_list=births_list))
pdf(file=file.path(FIGURES_FOLDER, "yearly_crs_incidence_timeseries.pdf"), w=8.5, h=11)
plots_per_page <- 5
groupings <- split(seq_along(plots), ceiling(seq_along(plots)/plots_per_page))
for (ii in seq_along(groupings)) {
  plots_subset <- plots[groupings[[ii]]]
  num_plots_subset <- length(plots_subset)
  if (num_plots_subset<plots_per_page) {
    for (jj in 1:(plots_per_page-num_plots_subset)) {
      plots_subset[[jj+num_plots_subset]] <- cowplot::ggdraw()
    }
  }
  print(cowplot::plot_grid(plotlist=plots_subset, ncol=1))
}
a <- dev.off()
```

Here, we plot the yearly CRS dynamics at the median $R_0$ for the states. 

```{r}
# It takes the state name and returns a time series of the yearly CRS burden with uncertainty bars. 
plot_yearly_crs_dynamics <- function(these_states, state_highlights, crs_incidence_list, births_list, scenarios, file_name, legend=TRUE, scale_factor) {
  state_yearly_incidence_list <- list()
  for (this_state in these_states) {
    state_crs_incidence <- crs_incidence_list[[this_state]]
    state_births <- births_list[[this_state]]
    
    # compute the yearly crs incidence (per 100,000 live births) in each simulation
    yearly_incidence_list <- list()
    for (this_scenario in scenarios) {
      crs_incidence <- state_crs_incidence %>% 
        filter(scenario==this_scenario) %>% 
        arrange(-desc(time)) %>% 
        pull(crs)
      yearly_incidence <- rep(0, floor(length(crs_incidence)/365))
      births <- state_births[, 2]
      for (jj in seq(1, length(crs_incidence)-1, 365)) {
        yearly_incidence[floor(jj/365)+1] <- sum(crs_incidence[jj:(jj+364)])/sum(births[jj:(jj+364)])*10^5
      }
      yearly_incidence_list[[which(this_scenario==scenarios)]] <- yearly_incidence 
    }
    state_yearly_incidence_list[[this_state]] <- do.call(cbind, yearly_incidence_list) %>% 
      as.data.frame() %>% 
      setNames(scenarios) %>% 
      mutate(year=seq(1, floor(length(crs_incidence)/365), 1)) %>% 
      select(year, dplyr::everything()) %>% 
      gather(key="scenario", value="crs", 2:ncol(.)) %>% 
      mutate(state=this_state, 
             color=ifelse(scenario==state_highlights[which(this_state==these_states)], 
                          "red", "grey"))
  }
  state_yearly_incidence <- do.call(rbind, state_yearly_incidence_list) %>% 
    as.data.frame() %>% 
    mutate(scenario=factor(scenario, scenarios), 
           state=factor(state, levels=these_states)) %>% 
    mutate(color=ifelse(scenario=="baseline", "black", color))
  
  p <- ggplot(state_yearly_incidence) + 
    geom_line(mapping=aes(x=year, y=crs, color=color, group=scenario)) + 
    geom_point(mapping=aes(x=year, y=crs, shape=scenario), size=0.5) +
    labs(x="Year", y="CRS cases/100,000 live births") + 
    scale_x_continuous(breaks=seq(0, 30, 5), expand=c(0, 0)) +
    scale_color_manual(breaks=c("red", "black", "grey"), 
                       values=c("red", "black", "lightgrey"), guide="none") + 
    scale_shape_manual(values=c(15, 16, 17, 18, 8, 3),
                       labels=c("No vaccination", 
                                "Routine only", 
                                "Routine + 1% increase", 
                                "Routine + 2.5% increase", 
                                "Routine + 5% increase", 
                                "Routine + 10% increase"), 
                       guide=guide_legend(override.aes = list(size=2.5))) +
    theme_bw() + 
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          panel.background=element_blank(), panel.border=element_blank(),
          axis.line = element_line(colour = "black")) + 
    theme(legend.position="bottom", legend.title=element_blank(), 
          legend.text=element_text(size=8), legend.key.size=unit(0.3, "cm")) + 
    theme(axis.text = element_text(size=8, color="black"), axis.title.y = element_text(size=8, color="black"), 
          axis.title.x=element_blank(), strip.text=element_text(size=8, color="black"), 
          strip.background=element_blank()) + 
    facet_rep_wrap(~state, ncol=2, scales="free_y")
  
  if (!legend) p <- p + theme(legend.position="none")
  pdf(file=file.path(FIGURES_FOLDER, file_name), w=8, h=11*scale_factor)
  print(p)
  a <- dev.off()
  return(p)
}
```

```{r}
# selected states
selected_yearly_crs_dynamics <- plot_yearly_crs_dynamics(these_states=c("Adamawa", "Kebbi", "Borno", "Jigawa"), 
                                                         state_highlights=as.character(c(0, 0.01, 0.025, 0.05)), 
                                                         crs_incidence_list=basic_scenarios_medR0_crs_incidence_list, 
                                                         births_list=births_list, 
                                                         scenarios=c("baseline", 0, 0.01, 0.025, 0.05, 0.1), 
                                                         file_name="yearly_crs_incidence_timeseries_pub.pdf", 
                                                         scale_factor=0.3)
selected_yearly_crs_dynamics

# all states
states <- c(unique(pull(ng2_pop, state)), "Nigeria")
state_highlights <- simple_dynamic_crs_burden_dist %>% 
  left_join(rename(baseline_crs_burden_dist, baseline_crs_burden=crs_burden), 
            by=c("state", "sim")) %>%
  group_by(scenario, state) %>% 
  summarise(mean_relative_burden=mean(crs_burden/baseline_crs_burden), .groups="keep") %>% 
  ungroup() %>% group_by(state) %>% 
  mutate(state=factor(state, levels=states)) %>% 
  arrange(-desc(state), -desc(scenario)) %>% 
  mutate(is_min_scenario=(scenario==scenario[min(which(mean_relative_burden<1))])) %>%
  filter(is_min_scenario) %>% 
  pull(scenario)

plots_per_page <- 16
groupings <- split(seq_along(states), ceiling(seq_along(states)/plots_per_page))
scale_factors <- unlist(lapply(groupings, length))/plots_per_page*0.9
scale_factors[length(groupings)] <- scale_factors[length(groupings)] + 0.075
for (ii in seq_along(groupings)) {
  all_yearly_crs_dynamics <- plot_yearly_crs_dynamics(these_states=states[groupings[[ii]]], 
                                                      state_highlights=as.character(state_highlights)[groupings[[ii]]], 
                                                      crs_incidence_list=basic_scenarios_medR0_crs_incidence_list, 
                                                      births_list=births_list, 
                                                      legend=(ii==length(groupings)), 
                                                      scenarios=c("baseline", 0, 0.01, 0.025, 0.05, 0.1), 
                                                      file_name=paste0("yearly_crs_incidence_timeseries_group", ii, ".pdf"), 
                                                      scale_factor=scale_factors[ii])
}

```

### Change in 30-year CRS burden

First, let's have a look at how 30-year CRS burden changes in each state following the introduction of RCV vaccination under different rates of improvement in routine coverage over time. 

We will start by showing whether there is an expected increase or decrease relative to the non-introduction scenario. 

```{r fig.height=4}
plot_mean_crs_burden_binary_change_heatmap <- function(dynamic_crs_burden_dist, baseline_crs_burden_dist, 
                                                       scenarios, file_name) {
  crs_burden_summary <- dynamic_crs_burden_dist %>% 
    left_join(rename(baseline_crs_burden_dist, baseline_crs_burden=crs_burden), 
                     by=c("state", "sim")) %>%
    mutate(change_crs_burden=crs_burden-baseline_crs_burden) %>% 
    group_by(scenario, state) %>% 
    summarise(mean_change_crs_burden=mean(change_crs_burden), .groups="keep") %>% 
    mutate(binary_change=ifelse(mean_change_crs_burden>0, "increase", "decrease")) %>% 
    mutate(scenario=factor(scenario, levels=scenarios)) %>% 
    mutate(state=ifelse(state=="Federal Capital Territory", "FCT", state))
  state_order <- sort(unique(pull(crs_burden_summary, state)))
  state_order <- c(state_order[which(state_order!="Nigeria")], "Nigeria")
  crs_burden_summary %<>% mutate(state=factor(state, levels=state_order))

  axis_text_colors <- c(rep("black", length(state_order)-1), "red")
  
  mean_crs_burden_binary_change_heatmap <- ggplot(data=crs_burden_summary) + 
    geom_tile(mapping=aes(x=state, y=scenario, fill=binary_change, color=binary_change), color="black") + 
    labs(y="Rate of improvement", fill="") +
    scale_y_discrete(labels=paste0(round(scenarios*100, 1), "%"), expand=c(0, 0)) +
    scale_x_discrete(drop=FALSE, expand=c(0, 0)) + 
    scale_fill_manual(breaks=c("increase", "decrease"), values=brewer.pal(9, "Set1")[c(1, 3)], labels=c("Increase", "Decrease")) +
        theme(axis.text.x=element_text(angle=90, size=9, hjust=1, vjust=0.5, color=axis_text_colors), axis.text=element_text(size=9, color="black"), 
          axis.title.x=element_blank(),
          axis.title.y=element_text(size=9, color="black"), 
          legend.position="right", legend.text=element_text(size=9, color="black"), legend.title=element_text(size=9, color="black")) +
    theme(panel.background=element_blank(), panel.grid=element_blank(), panel.border=element_rect(color="black", fill=NA))
  
  return(mean_crs_burden_binary_change_heatmap)
}
mean_crs_burden_binary_change_heatmap <- plot_mean_crs_burden_binary_change_heatmap(dynamic_crs_burden_dist=simple_dynamic_crs_burden_dist, 
                                                                                    baseline_crs_burden_dist=baseline_crs_burden_dist, 
                                                                                    scenarios=scenarios, 
                                                                                    file_name="mean_crs_burden_binary_change_heatmap.pdf")
mean_crs_burden_binary_change_heatmap
```

Let's have a look at how CRS burden is expected to quantitatively change. 

```{r fig.height=4, fig.width=8}
plot_mean_crs_burden_change_heatmap <- function(ng1_polygons=ng1, dynamic_crs_burden_dist, baseline_crs_burden_dist, scenarios, file_name) {
  crs_burden_summary <- dynamic_crs_burden_dist %>% 
    left_join(rename(baseline_crs_burden_dist, baseline_crs_burden=crs_burden), 
                     by=c("state", "sim")) %>%
    mutate(change_crs_burden=crs_burden-baseline_crs_burden) %>% 
    group_by(scenario, state) %>% 
    summarise(mean_percent_change_crs_burden=mean(change_crs_burden/baseline_crs_burden)*100, 
              mean_relative_burden=mean(crs_burden/baseline_crs_burden), .groups="keep") %>% 
    mutate(scenario=factor(scenario, levels=scenarios)) %>% 
    ungroup() %>% group_by(state) %>% 
    arrange(-desc(state), -desc(scenario)) %>% 
    mutate(is_min_scenario=(scenario==scenario[min(which(mean_relative_burden<1))])) %>% 
    mutate(state=ifelse(state=="Federal Capital Territory", "FCT", state))
  state_order <- sort(unique(pull(crs_burden_summary, state)))
  state_order <- c(state_order[which(state_order!="Nigeria")], "Nigeria")
  crs_burden_summary %<>% mutate(state=factor(state, levels=state_order))

  # heat map
  axis_text_colors <- c(rep("black", length(state_order)-1), "red")
  mean_crs_burden_change_heatmap <- ggplot() + 
    geom_tile(data=crs_burden_summary, 
              mapping=aes(x=state, y=scenario, fill=mean_relative_burden, color=mean_relative_burden)) + 
    geom_point(data=filter(crs_burden_summary, is_min_scenario),
               mapping=aes(x=state, y=scenario), shape=18) + 
    scale_fill_gradientn(colours=rev(brewer.pal(11, "RdBu")), breaks=seq(0, 2, 0.5), 
                         limits=c(0, 2.1), name="Relative\nburden") + 
    scale_color_gradientn(colours=rev(brewer.pal(11, "RdBu")), breaks=seq(0, 2, 0.5), 
                          limits=c(0, 2.1), guide="none") + 
    labs(y="Rate of improvement", fill="Relative\nburden") +
    scale_y_discrete(labels=paste0(round(scenarios*100, 1), "%"), expand=c(0, 0)) +
    scale_x_discrete(drop=FALSE, expand=c(0, 0)) + 
    guides(fill=guide_colorbar(title.position="top", title.hjust=0.5, 
                             ticks.colour="black", frame.colour="black")) + 
    theme(axis.text.x=element_text(angle=90, size=8, hjust=1, vjust=0.5, color=axis_text_colors), axis.text=element_text(size=8, color="black"), 
          axis.title.x=element_blank(),
          axis.title.y=element_text(size=8, color="black"), 
          legend.position="right", legend.text=element_text(size=8, color="black"), legend.title=element_text(size=8, color="black"), 
          legend.key.size=unit(0.3, "cm"), legend.key.width=unit(0.5, "cm"), legend.margin=unit(c(0, 0, 0, 0), "cm")) +
    theme(panel.background=element_blank(), panel.grid=element_blank(), panel.border=element_rect(color="black", fill=NA))
  
  # geographical map
  ng1_polygons %<>% mutate(state=ifelse(state=="Federal Capital Territory", "FCT", state))
  min_scenario_map_data <- crs_burden_summary %>% 
    ungroup() %>% 
    filter(is_min_scenario, state!="Nigeria") %>% 
    right_join(ng1_polygons, by="state")
    
  color_scale <- rev(brewer.pal(5, "RdYlBu"))
  color_scale <- brewer.pal(5, "Reds")
  min_scenario_map <- ggplot(data=min_scenario_map_data) + 
    geom_sf(mapping=aes(geometry=geometry, fill=scenario), color="black", size=0.3) + 
    scale_fill_manual(breaks=rev(c(0, 0.01, 0.025, 0.05, 0.10)), values=rev(color_scale), 
                      labels=rev(paste0(round(scenarios*100, 1), "%")), drop=FALSE) + 
    theme_bw() + 
    guides(fill=guide_legend(title.position="top", title.hjust=0.5, 
                             ticks.colour="black", frame.colour="black")) + 
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          panel.background=element_blank(), panel.border=element_blank(), 
          axis.line=element_blank(), axis.ticks=element_blank(), axis.text=element_blank(), 
          axis.title=element_blank()) + 
    theme(legend.position=c(0.9, 0.2), legend.key.size=unit(0.3, "cm"), legend.key.width=unit(0.5, "cm"), 
          legend.text=element_text(size=8), legend.title=element_blank(), 
          legend.box.background=element_blank(), legend.background=element_blank()) + 
    theme(plot.margin=unit(c(0, 0.5, 0, 0), "cm"))
  
  pAB <- cowplot::plot_grid(mean_crs_burden_change_heatmap, min_scenario_map, nrow=1, rel_widths=c(1, 0.4), 
                            labels=c("(a)", "(b)"), label_size=10, label_fontface="plain", label_fontfamily="serif")
  return(pAB)
}
mean_crs_burden_change_heatmap <- plot_mean_crs_burden_change_heatmap(dynamic_crs_burden_dist=simple_dynamic_crs_burden_dist, 
                                                                      baseline_crs_burden_dist=baseline_crs_burden_dist, 
                                                                      scenarios=c(0, 0.01, 0.025, 0.05, 0.1), 
                                                                      file_name="mean_crs_burden_change_heatmap.pdf")
mean_crs_burden_change_heatmap
```

Let's combine a couple of the plots...

```{r}
mean_crs_burden_change_heatmap <- plot_mean_crs_burden_change_heatmap(dynamic_crs_burden_dist=simple_dynamic_crs_burden_dist, 
                                                                      baseline_crs_burden_dist=baseline_crs_burden_dist, 
                                                                      scenarios=c(0, 0.01, 0.025, 0.05, 0.1), 
                                                                      file_name="mean_crs_burden_change_heatmap.pdf")
selected_yearly_crs_dynamics <- plot_yearly_crs_dynamics(these_states=c("Adamawa", "Kebbi", "Borno", "Jigawa"), 
                                                         state_highlights=as.character(c(0, 0.01, 0.025, 0.05)), 
                                                         crs_incidence_list=basic_scenarios_medR0_crs_incidence_list, 
                                                         births_list=births_list, 
                                                         scenarios=c("baseline", 0, 0.01, 0.025, 0.05, 0.1), 
                                                         file_name="yearly_crs_incidence_timeseries_pub.pdf", 
                                                         scale_factor=0.3)
pAB <- cowplot::plot_grid(mean_crs_burden_change_heatmap, 
                          selected_yearly_crs_dynamics, 
                          labels=c("", "(c)"), label_size=10, label_fontface="plain", label_fontfamily="serif", 
                          rel_heights=c(0.7, 1), 
                          ncol=1)
scale_factor <- 0.45
pdf(file=file.path(FIGURES_FOLDER, "effect_improvements_in_coverage.pdf"), w=8, h=11*scale_factor)
print(pAB)
a <- dev.off()
```

### Relative temporal dynamics

Let's start by computing the yearly CRS incidence in each scenario for each state. 

```{r}
# compute the yearly crs incidence (per 100,000 live births) in each simulation
scenarios <- c("baseline", 0, 0.01, 0.025, 0.05, 0.1) 
states <- c(unique(pull(ng2_pop, state)), "Nigeria")

yearly_incidence_summary_list <- list()
for (this_state in states) {
  state_crs_incidence_list <- all_basic_scenarios_crs_incidence_list[[this_state]]
  baseline_state_crs_incidence <- baseline_crs_incidence_list[[this_state]]
  state_births <- births_list[[this_state]]
  state_crs_incidence_list <- c(list(baseline_state_crs_incidence), state_crs_incidence_list)
  
  state_yearly_incidence_summary_list <- list()
  for (this_scenario in seq_along(scenarios)) {
    state_crs_incidence <- state_crs_incidence_list[[this_scenario]]
    
    yearly_incidence_list <- list()
    yearly_incidence <- rep(0, floor(nrow(state_crs_incidence)/365))
    for (ii in 1:(ncol(state_crs_incidence)-1)) {
      crs_incidence <- state_crs_incidence[, 1+ii]
      births <- state_births[, 2]
      for (jj in seq(1, length(crs_incidence)-1, 365)) {
        yearly_incidence[floor(jj/365)+1] <- sum(crs_incidence[jj:(jj+364)])/sum(births[jj:(jj+364)])*10^5
      }
      yearly_incidence_list[[ii]] <- yearly_incidence
    }
    state_yearly_incidence <- do.call(cbind, yearly_incidence_list) %>% 
      as.data.frame() %>% 
      setNames(1:ncol(.)) %>% 
      mutate(year=1:nrow(.)) %>% 
      dplyr::select(year, dplyr::everything()) %>% 
      gather(key="sim", value="crs_burden", 2:ncol(.)) %>% 
      mutate(scenario=scenarios[this_scenario], state=this_state)
    state_yearly_incidence_summary_list[[this_scenario]] <- state_yearly_incidence
  }
  yearly_incidence_summary_list[[this_state]] <- do.call(rbind, state_yearly_incidence_summary_list)
}
yearly_incidence_summary <- do.call(rbind, yearly_incidence_summary_list)
row.names(yearly_incidence_summary) <- NULL
```

Let's see the distribution in the relative yearly CRS incidence between the various vaccination scenarios and the non-vaccination scenario. 

```{r}
plot_expected_yearly_incidence_distributions <- function(these_states, scenarios, yearly_incidence_summary, 
                                                         dynamic_crs_burden_dist, baseline_crs_burden_dist, 
                                                         file_name) {
  # long-term crs burden summary
  crs_burden_summary <- dynamic_crs_burden_dist %>% 
    left_join(rename(baseline_crs_burden_dist, baseline_crs_burden=crs_burden), 
                     by=c("state", "sim")) %>%
    mutate(change_crs_burden=crs_burden-baseline_crs_burden, 
           relative_crs_burden=crs_burden/baseline_crs_burden, 
           percent_change_crs_burden=(crs_burden-baseline_crs_burden)/baseline_crs_burden*100) %>% 
    mutate(scenario=factor(scenario, levels=scenarios))
  
  # yearly incidence summmary
  baseline_yearly_incidence_summary <- yearly_incidence_summary %>% 
    filter(scenario=="baseline") %>% 
    select(-scenario) %>% 
    rename(baseline_crs_burden=crs_burden)
  yearly_incidence_summary <- yearly_incidence_summary %>% 
    filter(scenario!="baseline") %>% 
    left_join(baseline_yearly_incidence_summary, by=c("year", "sim", "state")) %>% 
    mutate(crs_burden_diff=crs_burden-baseline_crs_burden) %>% 
    mutate(above=crs_burden_diff>0) %>% 
    group_by(state, scenario, sim) %>% 
    summarise(num_yrs_above=sum(above), 
              rel_above=ifelse(num_yrs_above==0, NA, mean(crs_burden_diff[above])), .groups="keep") %>% 
    as.data.frame() %>% 
    mutate(scenario=factor(scenario, levels=scenarios), 
           sim=as.numeric(sim))
  
  # combined simulation summary
  combined_crs_summary <- yearly_incidence_summary %>% 
    left_join(crs_burden_summary, by=c("state", "scenario", "sim")) %>% 
    filter(state %in% these_states) %>% 
    select(state, scenario, sim, change_crs_burden, num_yrs_above, rel_above) %>% 
    gather(key="metric", value="value", c("change_crs_burden", "num_yrs_above", "rel_above")) %>% 
    mutate(metric=factor(metric, levels=c("change_crs_burden", "num_yrs_above", "rel_above"))) %>% 
    group_by(state, scenario, metric) %>% 
    summarise(median=median(value, na.rm=TRUE), 
              l50=quantile(value, 0.25, na.rm=TRUE), 
              u50=quantile(value, 0.75, na.rm=TRUE), .groups="keep")
  these_states[which(these_states=="Federal Capital Territory")] <- "FCT"
  combined_crs_summary %<>%
    mutate(state=ifelse(state=="Federal Capital Territory", "FCT", state)) %>%
    mutate(state=factor(state, levels=these_states))

  dummy_variable_for_axes <- data.frame(state=pull(combined_crs_summary, state)[1], 
                                        scenario=pull(combined_crs_summary, scenario)[1], 
                                        metric=c("change_crs_burden", "num_yrs_above", "rel_above"), 
                                        median=c(0, 0, 0), 
                                        l50=c(-100, 0, 0), 
                                        u50=c(20, 30, 40))

  metric_label <- c(change_crs_burden="Change in 30-year CRS burden", 
                    num_yrs_above="Number years above baseline", 
                    rel_above="Relative yearly CRS incidence")
  p <- ggplot(data=combined_crs_summary) + 
    geom_errorbar(dummy_variable_for_axes, mapping=aes(xmin=l50, xmax=u50, y=scenario), width=0, color="lightgrey") + 
    geom_vline(xintercept=0, linetype="dashed", color="black") + 
    geom_errorbar(mapping=aes(xmin=l50, xmax=u50, y=scenario), color="darkred", width=0, size=0.7) + 
    geom_point(mapping=aes(x=median, y=scenario), color="black", shape=21, fill="white", size=1.2) +
    scale_y_discrete(labels=paste0(round(scenarios*100, 1), "%"), name="Rate of Improvement") + 
    facet_grid(vars(state), vars(metric), scale="free", labeller=labeller(metric=metric_label)) + 
    theme_bw() + 
    theme(axis.title.x=element_blank()) + 
    theme(panel.spacing.y = unit(0, "lines"), strip.background=element_rect(fill="white")) + 
    theme(panel.grid=element_blank(), panel.grid.major.y=element_line(color="lightgrey"), 
          axis.text=element_text(color="black", size=7), 
          strip.text=element_text(color="black", size=8), 
          axis.title=element_text(color="black", size=8))
    
  
  scale_factor <- length(these_states)/14*0.9
  pdf(file=file.path(FIGURES_FOLDER, file_name), w=8, h=11*scale_factor)
  print(p)
  a <- dev.off()
}
```

```{r}
all_states <- c(unique(pull(ng2_pop, state)), "Nigeria")
plots_per_page <- 14
groupings <- split(seq_along(all_states), ceiling(seq_along(all_states)/plots_per_page))
for (ii in seq_along(groupings)) {
  p <- plot_expected_yearly_incidence_distributions(these_states=all_states[groupings[[ii]]], 
                                                    scenarios=c(0, 0.01, 0.025, 0.05, 0.1),
                                                    yearly_incidence_summary=yearly_incidence_summary, 
                                                    dynamic_crs_burden_dist=simple_dynamic_crs_burden_dist, 
                                                    baseline_crs_burden_dist=baseline_crs_burden_dist,
                                                    file_name=paste0("yearly_crs_incidence_dynamics_", "group", ii))
}
```


## CRS burden: advanced dynamic analysis

In this section, we focus on the dynamic analysis of simulations where routine infant vaccination coverage changes over time alongside regular supplemental immunization campaigns. 

### Yearly CRS incidence (time-series)

Here, we plot the yearly CRS dynamics at the median $R_0$ for all states. 

```{r}
# It takes the state name and returns a time series of the yearly CRS burden with uncertainty bars. 
plot_yearly_crs_dynamics <- function(these_states, state_highlights, crs_incidence_list, births_list, scenarios, file_name, legend=TRUE, scale_factor) {
  state_yearly_incidence_list <- list()
  for (this_state in these_states) {
    state_crs_incidence <- crs_incidence_list[[this_state]]
    state_births <- births_list[[this_state]]
    
    # compute the yearly crs incidence (per 100,000 live births) in each simulation
    yearly_incidence_list <- list()
    for (this_scenario in scenarios) {
      crs_incidence <- state_crs_incidence %>% 
        filter(scenario==this_scenario) %>% 
        arrange(-desc(time)) %>% 
        pull(crs)
      yearly_incidence <- rep(0, floor(length(crs_incidence)/365))
      births <- state_births[, 2]
      for (jj in seq(1, length(crs_incidence)-1, 365)) {
        yearly_incidence[floor(jj/365)+1] <- sum(crs_incidence[jj:(jj+364)])/sum(births[jj:(jj+364)])*10^5
      }
      yearly_incidence_list[[which(this_scenario==scenarios)]] <- yearly_incidence 
    }
    state_yearly_incidence_list[[this_state]] <- do.call(cbind, yearly_incidence_list) %>% 
      as.data.frame() %>% 
      setNames(scenarios) %>% 
      mutate(year=seq(1, floor(length(crs_incidence)/365), 1)) %>% 
      select(year, dplyr::everything()) %>% 
      gather(key="scenario", value="crs", 2:ncol(.)) %>% 
      mutate(state=this_state, 
             color=ifelse(scenario==state_highlights[which(this_state==these_states)], 
                          "red", "grey"))
  }
  state_yearly_incidence <- do.call(rbind, state_yearly_incidence_list) %>% 
    as.data.frame() %>% 
    mutate(scenario=factor(scenario, scenarios), 
           state=factor(state, levels=these_states)) %>% 
    mutate(color=ifelse(scenario=="baseline", "black", color))
  
  p <- ggplot(state_yearly_incidence) + 
    geom_line(mapping=aes(x=year, y=crs, color=color, group=scenario)) + 
    geom_point(mapping=aes(x=year, y=crs, shape=scenario), size=0.5) +
    labs(x="Year", y="CRS cases/100,000 live births") + 
    scale_x_continuous(breaks=seq(0, 30, 5), expand=c(0, 0)) +
    scale_color_manual(breaks=c("red", "black", "grey"), 
                       values=c("red", "black", "lightgrey"), guide="none") + 
    scale_shape_manual(values=c(15, 16, 17, 18, 8, 3),
                       labels=c("No vaccination", 
                                "Routine + SIA", 
                                "Routine + SIA + 1% increase", 
                                "Routine + SIA + 2.5% increase", 
                                "Routine + SIA + 5% increase", 
                                "Routine + SIA + 10% increase"), 
                       guide=guide_legend(override.aes = list(size=2.5))) +
    theme_bw() + 
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          panel.background=element_blank(), panel.border=element_blank(),
          axis.line = element_line(colour = "black")) + 
    theme(legend.position="bottom", legend.title=element_blank(), 
          legend.text=element_text(size=8), legend.key.size=unit(0.5, "cm")) + 
    theme(axis.text = element_text(size=9), axis.title.y = element_text(size=9), 
          axis.title.x=element_blank(), strip.text=element_text(size=9), 
          strip.background=element_blank()) + 
    facet_rep_wrap(~state, ncol=2, scales="free_y")
  
  if (!legend) p <- p + theme(legend.position="none")
  pdf(file=file.path(FIGURES_FOLDER, file_name), w=8, h=11*scale_factor)
  print(p)
  a <- dev.off()
  return(p)
}
```

```{r}
# all states
states <- c(unique(pull(ng2_pop, state)), "Nigeria")
state_highlights <- simple_dynamic_crs_burden_dist %>% 
  left_join(rename(baseline_crs_burden_dist, baseline_crs_burden=crs_burden), 
            by=c("state", "sim")) %>%
  group_by(scenario, state) %>% 
  summarise(mean_relative_burden=mean(crs_burden/baseline_crs_burden), .groups="keep") %>% 
  ungroup() %>% group_by(state) %>% 
  mutate(state=factor(state, levels=states)) %>% 
  arrange(-desc(state), -desc(scenario)) %>% 
  mutate(is_min_scenario=(scenario==scenario[min(which(mean_relative_burden<1))])) %>%
  filter(is_min_scenario) %>% 
  pull(scenario)

plots_per_page <- 16
groupings <- split(seq_along(states), ceiling(seq_along(states)/plots_per_page))
scale_factors <- unlist(lapply(groupings, length))/plots_per_page*0.9
scale_factors[length(groupings)] <- scale_factors[length(groupings)] + 0.075
for (ii in seq_along(groupings)) {
  all_yearly_crs_dynamics <- plot_yearly_crs_dynamics(these_states=states[groupings[[ii]]], 
                                                      state_highlights=as.character(state_highlights)[groupings[[ii]]], 
                                                      crs_incidence_list=advanced_scenarios_sia2_medR0_crs_incidence_list, 
                                                      births_list=births_list, 
                                                      legend=(ii==length(groupings)), 
                                                      scenarios=c("baseline", 0, 0.01, 0.025, 0.05, 0.1), 
                                                      file_name=paste0("advanced_sia2_yearly_crs_incidence_timeseries_group", ii, ".pdf"), 
                                                      scale_factor=scale_factors[ii])
}
```

Here, we plot the yearly CRS dynamics at the median $R_0$ across a range of different scenarios. 

```{r}
# It takes the state name and returns a time series of the yearly CRS burden with uncertainty bars. 
combined_yearly_crs_dynamics <- function(these_states, crs_incidence_list, births_list, crs_burden_dist, scenarios, file_name, scale_factor, legend=TRUE) {
  # yearly crs incidence (per 100,000 live births) for each scenario
  state_yearly_incidence_list <- list()
  for (this_state in these_states) {
    state_crs_incidence <- crs_incidence_list[[this_state]]
    state_births <- births_list[[this_state]]
    yearly_incidence_list <- list()
    for (ii in 1:nrow(scenarios)) {
      this_sia <- scenarios[ii, "sia"]
      this_scenario <- scenarios[ii, "scenario"]
      crs_incidence <- state_crs_incidence %>% 
        filter(scenario==this_scenario, sia==this_sia) %>% 
        arrange(-desc(time)) %>% 
        pull(crs)
      yearly_incidence <- rep(0, floor(length(crs_incidence)/365))
      births <- state_births[, 2]
      for (jj in seq(1, length(crs_incidence)-1, 365)) {
        yearly_incidence[floor(jj/365)+1] <- sum(crs_incidence[jj:(jj+364)])/sum(births[jj:(jj+364)])*10^5
      }
      yearly_incidence_list[[ii]] <- yearly_incidence 
    }
    state_yearly_incidence_list[[this_state]] <- do.call(cbind, yearly_incidence_list) %>% 
      as.data.frame() %>% 
      setNames(1:nrow(scenarios)) %>% 
      mutate(year=seq(1, floor(length(crs_incidence)/365), 1)) %>% 
      select(year, dplyr::everything()) %>% 
      gather(key="scenario", value="crs", 2:ncol(.)) %>% 
      mutate(sia=scenarios[as.numeric(scenario), "sia"], 
             scenario=scenarios[as.numeric(scenario), "scenario"]) %>% 
      mutate(state=this_state)
  }
  state_yearly_incidence <- do.call(rbind, state_yearly_incidence_list) %>% 
    as.data.frame() %>% 
    transmute(state=state, scenario=paste0(sia, "_", scenario), year=year, crs=crs) %>% 
    mutate(scenario=factor(scenario, paste0(scenarios[, "sia"], "_", scenarios[, "scenario"])),
           state=factor(state, levels=these_states))
  
  # long-term CRS burden
  selected_crs_burden_dist <- crs_burden_dist %>% 
    mutate(scenario=paste0(sia, "_", scenario)) %>%
    filter(scenario %in% paste0(scenarios[, "sia"], "_", scenarios[, "scenario"])) %>%
    group_by(state, scenario) %>% 
    summarise(med=median(crs_burden), 
              l25=quantile(crs_burden, 0.25), 
              u75=quantile(crs_burden, 0.75), .groups="keep") %>%
    as.data.frame() %>%
    filter(state %in% these_states) %>% 
    mutate(scenario=factor(scenario, paste0(scenarios[, "sia"], "_", scenarios[, "scenario"])),
           state=factor(state, levels=these_states))
    
  # y limits 
  state_y_limits <- select(state_yearly_incidence, state, crs) %>% 
    rbind(., transmute(selected_crs_burden_dist, state=state, crs=u75)) %>% 
    group_by(state) %>% 
    summarise(max_crs=max(crs)) %>% 
    mutate(state=factor(state, these_states)) %>% 
    arrange(-desc(state)) %>% 
    mutate(max_crs=plyr::round_any(max_crs, 5, f = ceiling)) %>% 
    pull(max_crs)
    
  # yearly CRS incidence plot
  colors <- c("grey", brewer.pal(12, "Paired")[c(5, 6, 3, 4, 2)])
  labels <- c("No vaccination", 
              "Routine only ", 
              "Routine only (2.5% increase)", 
              "Routine + starting campaign", 
              "Routine (2.5% increase) + starting campaign",
              "Routine (2.5% increase) + starting campaign + regular SIAs")
  dummy_p1_y_limits <- data.frame(state=factor(these_states, levels=these_states),  
                                  x=1, y=state_y_limits)
  p1_state_labels <- paste0("                                          ", these_states)
  names(p1_state_labels) <- these_states
  p1 <- ggplot(state_yearly_incidence) + 
    geom_line(mapping=aes(x=year, y=crs, group=scenario, color=scenario)) + 
    geom_point(mapping=aes(x=year, y=crs, shape=scenario), size=0.7) +
    geom_point(data=dummy_p1_y_limits, mapping=aes(x=x, y=y), color="white") + 
    labs(x="Year", y="CRS cases/100,000 live births") + 
    scale_x_continuous(breaks=seq(0, 30, 5), expand=c(0, 0)) +
    scale_color_manual(breaks=paste0(scenarios[, "sia"], "_", scenarios[, "scenario"]),
                       values=colors, 
                       labels=labels, 
                       guide=guide_legend(override.aes = list(size=1))) +
    scale_shape_manual(values=c(15, 16, 17, 18, 8, 3),
                       breaks=paste0(scenarios[, "sia"], "_", scenarios[, "scenario"]),
                       labels=labels) +
    theme_bw() + 
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          panel.background=element_blank(), panel.border=element_blank(),
          axis.line = element_line(colour = "black"), 
          panel.spacing.y=unit(-0.5, "cm")) + 
    theme(legend.position="bottom", legend.title=element_blank(), 
          legend.text=element_text(size=8), legend.key.size=unit(0.5, "cm")) + 
    theme(axis.text = element_text(size=9, color="black"), axis.title.y = element_text(size=9, color="black"), 
          axis.title.x=element_blank(), strip.text=element_text(size=9, vjust=-1), 
          strip.background=element_blank()) + 
    facet_rep_wrap(~state, ncol=1, scales="free_y", labeller=labeller(state=p1_state_labels))
  p1_legend <- get_legend(p1)
  p1 <- p1 + theme(legend.position="none")
  
  # long-term CRS burden plot
  dummy_state_labels <- rep("", length(these_states))
  names(dummy_state_labels) <- these_states
    dummy_p1_y_limits <- data.frame(state=factor(these_states, levels=these_states),  
                                  x=paste0(scenarios[1, "sia"], "_", scenarios[1, "scenario"]), 
                                  y=state_y_limits) %>% 
      mutate(x=factor(x, paste0(scenarios[, "sia"], "_", scenarios[, "scenario"])))
  p2 <- ggplot(data=selected_crs_burden_dist) + 
    geom_point(data=dummy_p1_y_limits, mapping=aes(x=x, y=y), color="white") + 
    geom_bar(mapping=aes(x=scenario, y=med, fill=scenario), stat="identity", color="black") + 
    geom_errorbar(mapping=aes(x=scenario, ymin=l25, ymax=u75), color="black", width=0.1) + 
    scale_x_discrete(labels=rep("", length(colors)), drop=FALSE) + 
    scale_y_continuous(expand=c(0, 0)) + 
    scale_fill_manual(breaks=paste0(scenarios[, "sia"], "_", scenarios[, "scenario"]),
                       values=colors, 
                       labels=rep("", length(colors)), 
                       guide="none") + 
    theme_bw() + 
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          panel.background=element_blank(), panel.border=element_blank(),
          axis.line = element_line(colour = "black"), 
          panel.spacing.y=unit(-0.5, "cm")) + 
    theme(axis.title.y = element_blank(), 
          axis.title.x=element_blank(), strip.text=element_text(size=9, color="black"), 
          strip.background=element_blank(), axis.text.y=element_blank()) + 
    facet_rep_wrap(~state, ncol=1, scales="free_y", labeller=labeller(state=dummy_state_labels))
  
  # combined plot
  p <- cowplot::plot_grid(p1, p2, nrow=1, rel_widths=c(1, 0.3))
  if (legend) p <- cowplot::plot_grid(p, p1_legend, ncol=1, rel_heights=c(1, 0.1))
  pdf(file=file.path(FIGURES_FOLDER, file_name), w=8, h=11*scale_factor)
  print(p)
  a <- dev.off()
  return(p)
}
```


```{r}
# organise simulation data
combined_crs_incidence_list <- purrr::map(1:length(advanced_scenarios_sia1_medR0_crs_incidence_list), 
                                          function(x) rbind(advanced_scenarios_sia1_medR0_crs_incidence_list[[x]], 
                                                            advanced_scenarios_sia2_medR0_crs_incidence_list[[x]]))
combined_crs_incidence_list <- purrr::map(1:length(combined_crs_incidence_list), 
                                          function(x) rbind(combined_crs_incidence_list[[x]], 
                                                            mutate(basic_scenarios_medR0_crs_incidence_list[[x]], 
                                                                   sia="sia0")))
names(combined_crs_incidence_list) <- names(advanced_scenarios_sia1_medR0_crs_incidence_list)
scenarios <- data.frame(sia=c("sia0", "sia0", "sia0", "sia2", "sia2", "sia1"),
                        scenario=c("baseline", "0", "0.025", "0", "0.025", "0.025"))
modified_simple_dynamics_crs_burden_dist <- simple_dynamic_crs_burden_dist %>% 
  transmute(state=state, crs_burden=crs_burden, sia="sia0", scenario=scenario, sim=sim)
all_crs_burden_dist <- do.call(rbind, list(modified_simple_dynamics_crs_burden_dist, 
                                           advanced_dynamic_sia1_crs_burden_dist, 
                                           advanced_dynamic_sia2_crs_burden_dist, 
                                           transmute(baseline_crs_burden_dist, 
                                                     state=state, crs_burden=crs_burden, sia="sia0", scenario="baseline", sim=sim)))

# plot for selected states
p <- combined_yearly_crs_dynamics(these_states=c("Adamawa", "Kebbi", "Borno", "Jigawa"), 
                                  crs_incidence_list=combined_crs_incidence_list, 
                                  births_list=births_list, 
                                  crs_burden_dist=all_crs_burden_dist,
                                  scenarios=scenarios, 
                                  file_name="advanced_combined_yearly_crs_incidence_timeseries_pub.pdf",
                                  scale_factor=0.45)
print(p)

# plot for all states
states <- c(unique(pull(ng2_pop, state)), "Nigeria")
plots_per_page <- 10
groupings <- split(seq_along(states), ceiling(seq_along(states)/plots_per_page))
scale_factors <- unlist(lapply(groupings, length))/plots_per_page*0.9
scale_factors[length(groupings)] <- scale_factors[length(groupings)] + 0.075
for (ii in seq_along(groupings)) {
  p <- combined_yearly_crs_dynamics(these_states=states[groupings[[ii]]],
                                    crs_incidence_list=combined_crs_incidence_list, 
                                    births_list=births_list, 
                                    crs_burden_dist=all_crs_burden_dist,
                                    scenarios=scenarios, 
                                    file_name=paste0("advanced_combined_yearly_crs_incidence_timeseries_group", ii, ".pdf"),
                                    scale_factor=scale_factors[ii], 
                                    legend=(ii==length(groupings)))
}

```

### Change in 30-year CRS burden

Let's compare the 30-year CRS burden under each scenario in each state to the no RCV introduction conditions.

```{r}
plot_complete_mean_crs_burden_change_heatmap <- function(crs_burden_dist, sia_types, scenarios, file_name) {
  baseline_crs_burden_dist <- crs_burden_dist %>% 
    filter(sia=="sia0", scenario=="baseline") %>% 
    transmute(state=state, sim=sim, baseline_crs_burden=crs_burden)
  crs_burden_summary <- crs_burden_dist %>% 
    filter(!(sia=="sia0" & scenario=="baseline")) %>% 
    left_join(baseline_crs_burden_dist, by=c("state", "sim")) %>%
    mutate(change_crs_burden=crs_burden-baseline_crs_burden) %>% 
    group_by(state, sia, scenario) %>% 
    summarise(mean_percent_change_crs_burden=mean(change_crs_burden/baseline_crs_burden)*100, 
              mean_relative_burden=mean(crs_burden/baseline_crs_burden), .groups="keep") %>% 
    mutate(scenario=factor(scenario, levels=scenarios), 
           sia=factor(sia, sia_types)) %>% 
    mutate(state=ifelse(state=="Federal Capital Territory", "FCT", state)) %>% 
    ungroup() %>% group_by(state, sia) %>%  
    arrange(-desc(state), -desc(sia), -desc(scenario)) %>% 
    mutate(is_min_scenario=(scenario==scenario[min(which(mean_relative_burden<1))])) 
  state_order <- sort(unique(pull(crs_burden_summary, state)))
  state_order <- c(state_order[which(state_order!="Nigeria")], "Nigeria")
  crs_burden_summary %<>% mutate(state=factor(state, levels=state_order))
  
  axis_text_colors <- c(rep("black", length(state_order)-1), "red")
  strip_labels <- c("Routine\n+ starting campaign\n+ regular SIAs", 
                    "Routine\n+ starting campaign", 
                    "Routine only")
  names(strip_labels) <- sia_types
  mean_crs_burden_change_heatmap <- ggplot() + 
    geom_tile(data=crs_burden_summary, 
              mapping=aes(x=state, y=scenario, fill=mean_relative_burden, color=mean_relative_burden)) + 
    geom_point(data=filter(crs_burden_summary, is_min_scenario),
               mapping=aes(x=state, y=scenario), shape=18) + 
    scale_fill_gradientn(colours=rev(brewer.pal(11, "RdBu")), breaks=seq(0, 2, 0.5), 
                         limits=c(0, 2.1), name="Relative\nburden") + 
    scale_color_gradientn(colours=rev(brewer.pal(11, "RdBu")), breaks=seq(0, 2, 0.5), 
                          limits=c(0, 2.1), guide="none") + 
    labs(y="Rate of improvement", fill="Relative\nburden") +
    scale_y_discrete(labels=paste0(round(scenarios*100, 1), "%"), expand=c(0, 0)) +
    scale_x_discrete(drop=FALSE, expand=c(0, 0)) + 
    guides(fill=guide_colorbar(title.position="top", title.hjust=0.5, 
                               ticks.colour="black", frame.colour="black")) + 
    theme(axis.text.x=element_text(angle=90, size=9, hjust=1, vjust=0.5, color=axis_text_colors), axis.text=element_text(size=9, color="black"), 
          axis.title.x=element_blank(),
          axis.title.y=element_text(size=9, color="black"), 
          legend.position="right", legend.text=element_text(size=9, color="black"), legend.title=element_text(size=9, color="black")) +
    theme(strip.background=element_blank(), panel.spacing.y = unit(0, "lines"), strip.placement="outside") + 
    theme(panel.background=element_blank(), panel.grid=element_blank(), panel.border=element_rect(color="black", fill=NA)) + 
    facet_grid(rows=vars(sia), labeller=labeller(sia=strip_labels))
  
  scale_factor <- 0.5
  pdf(file=file.path(FIGURES_FOLDER, file_name), w=8, h=11*scale_factor)
  print(mean_crs_burden_change_heatmap)
  a <- dev.off()
  return(mean_crs_burden_change_heatmap)
}
complete_mean_crs_burden_change_heatmap <- plot_complete_mean_crs_burden_change_heatmap(crs_burden_dist=all_crs_burden_dist,
                                                                                        sia_types=c("sia1", "sia2", "sia0"),
                                                                                        scenarios=c(0, 0.01, 0.025, 0.05, 0.1),
                                                                                        file_name="complete_mean_crs_burden_change_heatmap.pdf")
complete_mean_crs_burden_change_heatmap
```