---
title: "Initialization of dynamic rubella models"
output:
  html_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par(mgp = c(2, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin3 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", mai = rep(.1, 4))
    else NULL
  }
)

knitr::opts_chunk$set(echo       = TRUE,
                      margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center")
```

## Links to local data sets

Here we use the following links to data.

```{r}
DATA_FOLDER <- file.path("..", "data")
POPULATION_DATA_FOLDER <- file.path(DATA_FOLDER, "population_data")
R0_ESTIMATION_FOLDER <- file.path("..", "R0_estimation")
MCMC_ESTIMATES_FOLDER <- file.path(R0_ESTIMATION_FOLDER, "MCMC_estimates")
PARAMETERS_FOLDER <- file.path("parameters")
INITIAL_CONDITIONS_FOLDER <- file.path("initial_conditions")

ng2_pop_path <- file.path(POPULATION_DATA_FOLDER, "ng2_pop.rds")
ng0_pop_path <- file.path(POPULATION_DATA_FOLDER, "ng0_pop.rds")
```

Change them accordingly if you want to run the script locally on your computer.

## Preambule

This document generates the initial conditions for the MSEIRV age-structured model used for the rubella simulations in each of the thirty-seven states and at the national level. The R Markdown document `nigeria_data_organization_exploration.Rmd` must be successfully run prior to this document because it is responsible for downloading and organizing the following data files: `ng2_pop.rds` (population age and sex structure at local government level) and `ng0_pop.rds` (population age and sex structure at the national level). The R Markdown document `R0_estimation.Rmd` must also be successfully run prior to this document to derive the estimates for the force of infection functions, contact structure and initial seropositivity.  

## Packages

The needed packages:

```{r message = FALSE}
library(magrittr)
library(tidyverse)
```

## Data

### Nigeria demographic data

Here, we read in some basic demographic data (age-structure/gender split) at the local government level and national level in Nigeria (see `nigeria_data_organization_exploration.Rmd` for details on how the data was collected).

```{r}
ng2_pop <- readRDS(ng2_pop_path)
ng0_pop <- readRDS(ng0_pop_path)
```

### MCMC fitting data

Next, we read in the MCMC fitting data from the study of the seroprevalence data (see `R0_estimation.Rmd` for details).

```{r}
ng1_mean_seropositivity_estimates <- readRDS(file.path(MCMC_ESTIMATES_FOLDER,
                                                       "MCMC_piecewise_lambda_fit_mean_seropositivity_estimates.rds"))
ng1_lambda_estimates <- readRDS(file.path(MCMC_ESTIMATES_FOLDER, 
                                          "MCMC_piecewise_lambda_fit_lambda_estimates.rds"))
contact_structure <- readRDS(file.path(MCMC_ESTIMATES_FOLDER,
                                       "MCMC_piecewise_lambda_fit_contact_structure.rds"))
ng0_mean_seropositivity_estimates <- readRDS(file.path(MCMC_ESTIMATES_FOLDER,
                                                       "ng0_MCMC_piecewise_lambda_fit_mean_seropositivity_estimates.rds"))
ng0_lambda_estimates <- readRDS(file.path(MCMC_ESTIMATES_FOLDER, 
                                          "ng0_MCMC_piecewise_lambda_fit_lambda_estimates.rds"))
```

## Initial conditions

In this section, we distribute the individuals in each age group across the ten epidemiological classes (i.e. $M, S, E, I, R, V, SP_1, SP_2, SP_3, EP$) for each state in Nigeria.

```{r}
# `initialize_pop` takes as arguments the state name string `this_state` and returns a named numeric vector of the initial conditions for that state. 
initialize_pop <- function(this_state, mean_seropositivity_estimates, 
                           lambda_estimates, contact_structure) {
  # retrieve initial age distribution in months
  if (this_state=="Nigeria") 
    initial_age_distribution <- readRDS(file=file.path(INITIAL_CONDITIONS_FOLDER, "initial_age_distribution_country.rds"))
  else 
    initial_age_distribution <- readRDS(file=file.path(INITIAL_CONDITIONS_FOLDER, "initial_age_distribution_list.rds"))[[this_state]]
  
  # retrieves the basic parameters
  parameters <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "basic_parameters.rds"))
  
  # retrieve seropositivity estimates for the state
  state_sero_pos_data <- mean_seropositivity_estimates[[this_state]]
  
  # retrieve force of infection data for the state (we use the mean force of infection)
  state_foi_data <- lambda_estimates[[this_state]] %>% colMeans()
  base_agegrps <- contact_structure$agegrps
  
  # distribute population across epidemiological classes
  classes <- c("M", "S", "E", "I", "R", "V", "1SP", "2SP", "3SP", "EP")
  init_conds_list <- list()
  for (ii in 1:nrow(initial_age_distribution)) {
    agegrp <- initial_age_distribution[ii, "agegrp"]
    agegrp_n <- initial_age_distribution[ii, "n"]
    agegrp_index <- ii
    
    # get seropositivity for age group
    seropos <- state_sero_pos_data[findInterval(agegrp/12, state_sero_pos_data$age-1), "seropos"]
    
    # get foi for age group
    foi <- state_foi_data[findInterval(agegrp/12, base_agegrps)]/365
    
    # set initial conditions
    THREE_MONTHS_TO_DAYS <- 90
    if (agegrp==0) {
      initR <- 0; initM <- (seropos*agegrp_n)
    } else {
      initR <- (seropos*agegrp_n); initM <- 0
    }
    initI <- (1-seropos)*agegrp_n*foi*(1/parameters$gamma[paste("gamma", ii, sep="")])
    initE <- (1-seropos)*agegrp_n*foi*(1/parameters$sigma[paste("sigma", ii, sep="")])
    initS <- agegrp_n - initR - initI - initE - initM
    initSP1 <- parameters$f[paste("f", ii, sep="")]*initS*(THREE_MONTHS_TO_DAYS)
    initSP2 <- parameters$f[paste("f", ii, sep="")]*initS*(THREE_MONTHS_TO_DAYS)
    initSP3 <- parameters$f[paste("f", ii, sep="")]*initS*(THREE_MONTHS_TO_DAYS)
    initS <- initS - initSP1 - initSP2 - initSP3
    initEP <- initSP1*foi*(1/parameters$sigma[paste("sigma", ii, sep="")])
    initE <- initE - initEP
    initV <- 0
    init_conds <- c(initM, initS, initE, initI, initR, initV, initSP1, initSP2, initSP3, initEP)
    names(init_conds) <- paste(classes, ii, sep="")
    init_conds_list[[ii]] <- init_conds
  }
  init_conds <- do.call(c, init_conds_list)
  init_conds <- init_conds[sapply(1:length(classes), 
                                  function(x) seq(x, nrow(initial_age_distribution)*length(classes), length(classes)))]
  return(init_conds)
}
```
 
### State-level

```{r}
states <- sort(unique(pull(ng2_pop, state)))
init_conds_list <- purrr::map(states, 
                              function(x) initialize_pop(this_state=x, 
                                                         mean_seropositivity_estimates=ng1_mean_seropositivity_estimates, 
                                                         lambda_estimates=ng1_lambda_estimates, 
                                                         contact_structure=contact_structure))
names(init_conds_list) <- states
saveRDS(init_conds_list, file=file.path(INITIAL_CONDITIONS_FOLDER, "initial_conditions_list.rds"))
```

### National-level

```{r}
init_conds <- initialize_pop(this_state="Nigeria", 
                             mean_seropositivity_estimates=list("Nigeria"=ng0_mean_seropositivity_estimates), 
                             lambda_estimates=list("Nigeria"=ng0_lambda_estimates), 
                             contact_structure=contact_structure)
saveRDS(init_conds, file=file.path(INITIAL_CONDITIONS_FOLDER, "initial_conditions_country.rds"))
```