---
title: "MSEIRV compartment model simulators"
output:
  html_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(
  margin1 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin2 = function(before, options, envir) {
    if (before) par(mgp = c(2, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  margin3 = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", mai = rep(.1, 4))
    else NULL
  }
)

knitr::opts_chunk$set(echo       = TRUE,
                      margin1    = TRUE,
                      fig.retina = 2,
                      fig.align  = "center")
```

## Links to local data sets

Here we use the following links to data.

```{r}
# directories 
DATA_FOLDER <- file.path("..", "data")
POPULATION_DATA_FOLDER <- file.path(DATA_FOLDER, "population_data")
VACCINATION_DATA_FOLDER <- file.path(DATA_FOLDER, "vaccination_data")
R0_ESTIMATION_FOLDER <- file.path("..", "R0_estimation")
MCMC_ESTIMATES_FOLDER <- file.path(R0_ESTIMATION_FOLDER, "MCMC_estimates")
PARAMETERS_FOLDER <- file.path("parameters")
INITIAL_CONDITIONS_FOLDER <- file.path("initial_conditions")
RCPP_FOLDER <- file.path("rcpp")

# derived data from `nigeria_data_organization_exploration.Rmd`
ng2_pop_path <- file.path(POPULATION_DATA_FOLDER, "ng2_pop.rds")

# estimates from `R0_estimation.Rmd`
r0_estimates_path <- file.path(MCMC_ESTIMATES_FOLDER,
                               "MCMC_piecewise_lambda_fit_R0_estimates.rds")
r0_estimates_country_path <- file.path(MCMC_ESTIMATES_FOLDER,
                               "ng0_MCMC_piecewise_lambda_fit_R0_estimates.rds")

# initial conditions from `initial_conditions.Rmd` and `parameterization.Rmd`
initial_conditions_path <- file.path(INITIAL_CONDITIONS_FOLDER, "initial_conditions_list.rds")
initial_age_distribution_path <- file.path(INITIAL_CONDITIONS_FOLDER, "initial_age_distribution_list.rds")
initial_conditions_country_path <- file.path(INITIAL_CONDITIONS_FOLDER, "initial_conditions_country.rds")
initial_age_distribution_country_path <- file.path(INITIAL_CONDITIONS_FOLDER, "initial_age_distribution_country.rds")

# local data 
ng1_mcv1_data_path <- file.path(VACCINATION_DATA_FOLDER, "ng1_mcv1_data.xlsx")
general_sia_scenario1_path <- file.path(VACCINATION_DATA_FOLDER, "summarised_general_sia_scenario1.xlsx")
general_sia_scenario2_path <- file.path(VACCINATION_DATA_FOLDER, "summarised_general_sia_scenario2.xlsx")
```

These are the paths we use for the simulation output. 

```{r}
ALL_OUTPUT_FOLDER <- "~/Desktop/output/"
VALIDATION_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "validation_simulations", "raw_simulations")
STATIC_RISK_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "static_risk_simulations", "raw_simulations")
LANDSCAPE_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "static_risk_simulations", "landscape_simulations")
DYNAMIC_RISK_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "dynamic_risk_simulations", "raw_simulations")
SUMMARY_STATIC_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "static_risk_simulations", "summary_simulations")
SUMMARY_DYNAMIC_SIMS_FOLDER <- file.path(ALL_OUTPUT_FOLDER, "dynamic_risk_simulations", "summary_simulations")
```

Change them accordingly if you want to run the script locally on your computer.

## Preambule

This document runs simulations of the MSEIRV age-structured model. The model itself is implemented in Cpp (`mspeirv_ode_model.cpp`). The following R Markdown documents must be successfully run prior to this document: `nigeria_data_organization_exploration.Rmd`, `R0_estimation.Rmd`, `initial_conditions.Rmd` and `parameterization.Rmd`. The vaccination data (`ng1_mcv1_data.xlsx`, `summarised_general_sia_scenario1.xlsx` and `summarised_general_sia_scenario2.xlsx`) are provided in the Github repository. 

## Packages

The needed packages:

```{r message = FALSE}
library(magrittr)
library(tidyverse)
library(Rcpp)
library(microbenchmark)
library(readxl)
library(nloptr)
```

## Rcpp code

Here, we compile the rcpp code that we need to run the simulation model. Details on the MSEIRV compartment model are provided in the supplementary text. 

```{r}
sourceCpp(file.path(RCPP_FOLDER, "mspeirv_ode_model.cpp"))
```

## Data

### Nigeria demographic data

Here, we read in some basic demographic data (age-structure/gender split) at the local government level in Nigeria (see `nigeria_data_organization_exploration.Rmd` for details on how the data was collected).

```{r}
ng2_pop <- readRDS(ng2_pop_path)
```

### Parameter/Initial conditions data 

Here, we read in all the parameter and initial conditions data that is required for the simulations. 

```{r}
# initial conditions data
initial_conditions <- readRDS(initial_conditions_path)
initial_conditions[["Nigeria"]] <- readRDS(initial_conditions_country_path)

# initial age distribution data
initial_age_distributions <- readRDS(initial_age_distribution_path)
initial_age_distributions[["Nigeria"]] <- readRDS(initial_age_distribution_country_path)

# parameter data
states <- sort(unique(pull(ng2_pop, state)))
basic_parameters_list <- list()
beta_matrices_list <- list()
omegas_list <- list()
for (this_state in states) {
  basic_parameters_list[[this_state]] <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "basic_parameters.rds"))
  beta_matrices_list[[this_state]] <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "beta_matrices.rds"))
  omegas_list[[this_state]] <- readRDS(file=file.path(PARAMETERS_FOLDER, this_state, "omegas.rds"))
}
basic_parameters_list[["Nigeria"]] <- readRDS(file=file.path(PARAMETERS_FOLDER, "Nigeria", "basic_parameters.rds"))
beta_matrices_list[["Nigeria"]] <- readRDS(file=file.path(PARAMETERS_FOLDER, "Nigeria", "beta_matrices.rds"))
omegas_list[["Nigeria"]] <- readRDS(file=file.path(PARAMETERS_FOLDER, "Nigeria", "omegas.rds"))                                   

# R0 estimation data
r0_estimates <- readRDS(r0_estimates_path)
r0_estimates_country <- readRDS(r0_estimates_country_path)
```

### Vaccination data

Here, we read in the state-level MCV1 coverage data for Nigeria and general supplemental immunization scenarios. 

```{r}
ng1_mcv1_data <- readxl::read_excel(ng1_mcv1_data_path) %>% 
  rbind(., data.frame(state="Nigeria", MCV1=54.0))
general_sia_scenarios <- list("sia1"=readxl::read_excel(path=general_sia_scenario1_path, na="NA"), 
                              "sia2"=readxl::read_excel(path=general_sia_scenario2_path, na="NA"))
```

## Helper functions

`reorder_initial_conditions` takes a named vector of initial conditions, the number of stages and the order of the classes, and then returns the initial conditions vector according to the specified orders. This is necessary for the Rcpp code. 
```{r}
reorder_initial_conditions <- function(initial_conditions, num_stages,  
                                       class_order=c("M", "S", "E", "I", "R", "V", "1SP", "2SP", "3SP", "EP")) {
  reordered_initial_conditions <- c()
  for (ii in class_order) {
    reordered_initial_conditions <- c(reordered_initial_conditions, 
                                      initial_conditions[paste(ii, 1:num_stages, sep="")])
  }
  return(reordered_initial_conditions)
}
```

## Static risk assessment simulations

Next, we generate the simulations necessary for static risk assessment in each of the states. 

Here, we construct a function `mseir_static_risk_sim` which takes as arguments a state, `this_state`, a time frame (in years) `tf`, `sim_type` (0 for without routine coverage and 1 for with routine coverage), the proportion of children immunized during routine coverage `routine_cov`and the simulation number (this is relevant for the uncertainty analysis), the external immigration rate and returns a data frame with the results of the deterministic ODE simulation. When `simplified_output` is `TRUE`, only the time series of the exposed-and-pregnant compartments are returned (otherwise all the simulation results are returned). 

```{r}
mseirv_static_sim  <- function(this_state, tf, sim_type, routine_cov, sim_num, ext_immi_rate, simplified_output=FALSE) {
  # get parameters
  basic_parameters <- basic_parameters_list[[this_state]]
  beta_matrix <- beta_matrices_list[[this_state]][[sim_num]]
  omega <- omegas_list[[this_state]][sim_num]
  
  omegas <- rep(omega, basic_parameters$num_stages)
  names(omegas) <- paste0("omega", 1:basic_parameters$num_stages, sep="")
  basic_parameters$beta <- beta_matrix
  basic_parameters$omega <- omegas
  parameters <- unlist(basic_parameters)
  names(parameters) <- gsub("^.*\\.","", names(parameters))
  parameters <- c(parameters, sim_type=sim_type, eps=ext_immi_rate)
  if (sim_type==1) {
    pv <- rep(routine_cov, tf) 
    names(pv) <- paste("pv", 1:tf, sep="")
    parameters <- c(parameters, pv)
  } else if (sim_type==2) {
    pv <- routine_cov
    names(pv) <- paste("pv", 1:(tf), sep="")
    parameters <- c(parameters, pv) 
  } 
  
  # get initial conditions
  initial_conditions <- initial_conditions[[this_state]]
  initial_conditions <- reorder_initial_conditions(initial_conditions, 
                                                   num_stages=parameters["num_stages"])
  classes <- c("M", "S", "E", "I", "R", "V", "1SP", "2SP", "3SP", "EP")
  initial_pop <- sapply(1:(length(initial_conditions)/length(classes)), function(x) sum(initial_conditions[paste0(classes, x)]))
  names(initial_pop) <- paste0("N", 1:length(initial_pop))
  parameters <- c(parameters, initial_pop)
  
  output <- deSolve::ode(y=initial_conditions, times=seq(0, tf*365-1, 1), func=get_ode_rates, parms=parameters, 
                         maxsteps=10^4, method="rk4") %>%
    as.data.frame()
  if (simplified_output) {
    output <- output[, c("time", paste("EP", 1:parameters["num_stages"], sep=""))]
  }
  return(output)
}
```

### No vaccination 

Let's run 30 years of simulations without any routine vaccination. Note: `sim_type=0` denotes simulations with no interventions. 

```{r}
no_vac_simulator <- function(this_state, sim_num) {
  tf <- 30
  ext_immi_rate <- 10/10^5/365 # 10 infected immigrant per 100,000 individuals per year
  output <- mseirv_static_sim(this_state=this_state, tf=tf, sim_type=0, sim_num=sim_num, 
                              ext_immi_rate=ext_immi_rate, simplified_output=TRUE)
  file_name <- paste(this_state, paste0("tf", tf), "noVac", paste0("simNum", sim_num), sep="_")
  saveRDS(output, file=file.path(STATIC_RISK_SIMS_FOLDER, file_name))
}
```

State level simulations...

```{r results="hide", eval=FALSE}
states <- unique(pull(ng2_pop, state))
sim_nums <- 1:200
input <- expand.grid(state=states, sim_num=sim_nums)
(parallel::mcmapply(no_vac_simulator, as.character(input[["state"]]), input[["sim_num"]], mc.cores=6))
```
NOTE: Full results (i.e. `simplified_output=FALSE`) are required for `sim_num=1`. 

National level simulations...

```{r results="hide", eval=FALSE}
sim_nums <- 1:200
input <- expand.grid(state="Nigeria", sim_num=sim_nums)
(parallel::mcmapply(no_vac_simulator, as.character(input[["state"]]), input[["sim_num"]], mc.cores=5))
```
NOTE: Full results (i.e. `simplified_output=FALSE`) are required for `sim_num=1`

### Vaccination at current MCV1 coverage 

Let's run 30 years of simulations with routine vaccination set to current MCV1 coverage. Note: `sim_type=1` denotes simulations with routine vaccination (no change in coverage over time). 

```{r}
vac_simulator <- function(this_state, sim_num) {
  tf <- 30
  routine_cov <- as.numeric(ng1_mcv1_data[which(ng1_mcv1_data$state==this_state), "MCV1"])
  ext_immi_rate <- 10/10^5/365 # 10 infected immigrant per 100,000 individuals per year
  output <- mseirv_static_sim(this_state=this_state, tf=tf, sim_type=1, sim_num=sim_num, 
                              routine_cov=routine_cov/100, ext_immi_rate=ext_immi_rate, simplified_output=TRUE)
  file_name <- paste(this_state, paste0("tf", tf), "vac", paste0("simNum", sim_num), sep="_")
  saveRDS(output, file=file.path(STATIC_RISK_SIMS_FOLDER, file_name))
}
```

State level simulations...

```{r results="hide", eval=FALSE}
states <- unique(pull(ng2_pop, state))
sim_nums <- 1:200
input <- expand.grid(state=states, sim_num=sim_nums)
(parallel::mcmapply(vac_simulator, as.character(input[["state"]]), input[["sim_num"]], mc.cores=5))
```

National level simulations...

```{r results="hide", eval=FALSE}
sim_nums <- 1:200
input <- expand.grid(state="Nigeria", sim_num=sim_nums)
(parallel::mcmapply(vac_simulator, as.character(input[["state"]]), input[["sim_num"]], mc.cores=5))
```

### Minimum necessary vaccination coverage 

Next, we perform some optimization to estimate the minimum routine vaccination coverage necessary to prevent an increase in 30-year CRS burden relative to pre-vaccination burden. We start by defining some functions that are necessary for the optimization procedure. First, we define a function that takes a simulation output and computes the 30-year CRS burden. 

```{r}
compute_crs_burden <- function(sim, sim0, params) {
  # compute crs incidence
  EP <- sim[, paste("EP", 1:params$num_stages, sep="")] %>% as.data.frame()
  sigma <- params$sigma
  c <- 0.65
  crs_incidence <- apply(EP, 1, function(x) sum(c*x*sigma))
  
  # compute births
  aN <- sapply(1:params$num_stages, function(x) rowSums(sim0[, 1+seq(x, ncol(sim0)-1, params$num_stages)]))
  initN <- aN[1, ]
  total_initN <- sum(initN)
  births <- rowSums(t(sapply(1:nrow(aN), function(x) aN[x, ]*initN/aN[x, ]*params$f*(1+params$db)^x)))
  
  # compute long-term CRS burden
  crs_burden <- sum(crs_incidence)/sum(births)*10^5
  
  return(crs_burden)
}
```

Next, we define the objective function and constraints of the optimization problem.

```{r}
# objective function
eval_f0 <- function(x, this_state, this_sim_num) {
  return(x)
}

# constraints 
eval_g0 <- function(x, this_state, this_sim_num) {
  basic_parameters <- basic_parameters_list[[this_state]]
  
  output0 <- readRDS(file.path(STATIC_RISK_SIMS_FOLDER, paste(this_state, "tf30", "noVac", 
                                                              paste0("simNum", this_sim_num), sep="_")))
  f0 <- compute_crs_burden(output0, output0, basic_parameters)
  
  outputV <- mseirv_static_sim(this_state=this_state, tf=30, sim_type=1, sim_num=this_sim_num, 
                               ext_immi_rate = 10/10^5/365, 
                               routine_cov=x, simplified_output=TRUE)
  fv <- compute_crs_burden(outputV, output0, basic_parameters)
  
  alpha <- 0.01
  return(fv-f0-alpha)
}
```

**State level analyses** \newline

Since we don't have the computational infrastructure to perform uncertainty analyses, we focus on finding the minimum necessary vaccination coverage at the median $R_0$ and the bounds of the $90\%$ credible interval estimated from the serological survey data. Here, we find the parameters that correspond to the median $R_0$ and the bounds of the $90\%$ credible interval for each state. 

```{r}
states <- sort(unique(pull(ng2_pop, state)))
state_median_r0_idxs <- rep(NA, length(states))
state_median_r0s <- rep(NA, length(states))
state_l90_r0_idxs <- rep(NA, length(states))
state_l90_r0s <- rep(NA, length(states))
state_u90_r0_idxs <- rep(NA, length(states))
state_u90_r0s <- rep(NA, length(states))
for (ii in seq_along(states)) {
  state_r0_estimates <- r0_estimates[, states[ii]]
  state_median_r0s[ii] <- sort(state_r0_estimates)[floor(length(state_r0_estimates)*0.5)]
  state_median_r0_idxs[ii] <- which(state_r0_estimates==sort(state_r0_estimates)[floor(length(state_r0_estimates)*0.5)])
  state_l90_r0s[ii] <- sort(state_r0_estimates)[floor(length(state_r0_estimates)*0.05)]
  state_l90_r0_idxs[ii] <- which(state_r0_estimates==sort(state_r0_estimates)[floor(length(state_r0_estimates)*0.05)])
  state_u90_r0s[ii] <- sort(state_r0_estimates)[floor(length(state_r0_estimates)*0.95)]
  state_u90_r0_idxs[ii] <- which(state_r0_estimates==sort(state_r0_estimates)[floor(length(state_r0_estimates)*0.95)])
}
names(state_median_r0_idxs) <- states
names(state_median_r0s) <- states
names(state_l90_r0_idxs) <- states
names(state_l90_r0s) <- states
names(state_u90_r0_idxs) <- states
names(state_u90_r0s) <- states
```

We need to perform run the simulations for the no vaccination scenario for the selected indices. Full results (i.e. `simplified_output=FALSE`) are required for these simulations. 

```{r eval=FALSE}
median_input <- data.frame(state=names(state_median_r0_idxs), 
                           sim_num=state_median_r0_idxs)
(parallel::mcmapply(no_vac_simulator, as.character(median_input[["state"]]), median_input[["sim_num"]], mc.cores=5))

l90_input <- data.frame(state=names(state_l90_r0_idxs), 
                        sim_num=state_l90_r0_idxs)
(parallel::mcmapply(no_vac_simulator, as.character(l90_input[["state"]]), l90_input[["sim_num"]], mc.cores=5))

u90_input <- data.frame(state=names(state_u90_r0_idxs), 
                        sim_num=state_u90_r0_idxs)
(parallel::mcmapply(no_vac_simulator, as.character(u90_input[["state"]]), u90_input[["sim_num"]], mc.cores=5))
```

Finally, let's perform the optimization. 

```{r results="hide", eval=FALSE}
# define the initial conditions
x0s <- rep(0.9, length(unique(pull(ng2_pop, state))))
names(x0s) <- sort(unique(pull(ng2_pop, state)))

# perform the optimization for each state
optimization_output_list <- list()
states <- sort(unique(pull(ng2_pop, state)))
r0_idxs_list <- list(state_median_r0_idxs, state_l90_r0_idxs, state_u90_r0_idxs)
r0s_list <- list(state_median_r0s, state_l90_r0s, state_u90_r0s)
file_name_prefix <- c("median", "l90", "u90")
for (jj in 1:length(r0_idxs_list)) {
  state_r0_idxs <- r0_idxs_list[[jj]]
  state_r0s <- r0s_list[[jj]]
  for (ii in states) {
    result <- nloptr(x0 = x0s[ii], 
                     eval_f = eval_f0,
                     lb = c(0),
                     ub = c(1), 
                     eval_g_ineq = eval_g0,
                     opts = list("algorithm"="NLOPT_LN_COBYLA",
                                 "xtol_rel"=0.001),
                     this_state=ii, this_sim_num=state_r0_idxs[ii])
    print(result)
    optimization_output_list[[ii]] <- result
  }
  saveRDS(optimization_output_list, file=file.path(SUMMARY_STATIC_SIMS_FOLDER, 
                                                   paste0(file_name_prefix[jj], "_r0_min_cov_optimization_output.rds")))
  saveRDS(state_r0_idxs, file=file.path(SUMMARY_STATIC_SIMS_FOLDER, 
                                        paste0(file_name_prefix[jj], "_r0_simNums_optimization.rds")))
  saveRDS(state_r0s, file=file.path(SUMMARY_STATIC_SIMS_FOLDER, 
                                    paste0(file_name_prefix[jj], "_r0_optimization.rds")))
}
```

**National level analyses** \newline

```{r}
median_r0 <- sort(r0_estimates_country)[floor(length(r0_estimates_country)*0.5)]
median_r0_idx <- which(r0_estimates_country==sort(r0_estimates_country)[floor(length(r0_estimates_country)*0.5)])
l90_r0 <- sort(r0_estimates_country)[floor(length(r0_estimates_country)*0.05)]
l90_r0_idx <- which(r0_estimates_country==sort(r0_estimates_country)[floor(length(r0_estimates_country)*0.05)])
u90_r0 <- sort(r0_estimates_country)[floor(length(r0_estimates_country)*0.95)]
u90_r0_idx <- which(r0_estimates_country==sort(r0_estimates_country)[floor(length(r0_estimates_country)*0.95)])
```

```{r eval=FALSE}
input <- data.frame(state="Nigeria", 
                    sim_num=c(median_r0_idx, l90_r0_idx, u90_r0_idx))
(parallel::mcmapply(no_vac_simulator, as.character(input[["state"]]), input[["sim_num"]], mc.cores=3))
```

```{eval=FALSE, results="hide"}
statistics <- c("median", "l90", "u90")
r0s <- c(median_r0, l90_r0, u90_r0)
names(r0s) <- statistics
r0_idxs <- c(median_r0_idx, l90_r0_idx, u90_r0_idx)
names(r0_idxs) <- statistics
optimization_output_list <- list()
for (jj in 1:length(statistics)) {
  r0 <- r0s[jj]
  r0_idx <- r0_idxs[jj]
  result <- nloptr(x0 = 0.9, 
                   eval_f = eval_f0,
                   lb = c(0),
                   ub = c(1), 
                   eval_g_ineq = eval_g0,
                   opts = list("algorithm"="NLOPT_LN_COBYLA",
                               "xtol_rel"=0.001),
                   this_state="Nigeria", this_sim_num=r0_idx)
  print(result)
  optimization_output_list[[jj]] <- result
}
saveRDS(optimization_output_list, file=file.path(SUMMARY_STATIC_SIMS_FOLDER, 
                                                 "ng0_r0_min_cov_optimization_output.rds"))
saveRDS(r0_idxs, file=file.path(SUMMARY_STATIC_SIMS_FOLDER, 
                                "ng0_r0_simNums_optimization.rds"))
saveRDS(r0s, file=file.path(SUMMARY_STATIC_SIMS_FOLDER, 
                            "ng0_r0_optimization.rds"))
```

### CRS burden landscape

First, we also compute the landscape for each individual state at its median $R_0$. 

Next, we perform the simulations. 

```{r}
landscape_simulator <- function(this_state, routine_cov, sim_num) {
  tf <- 30
  ext_immi_rate <- 10/10^5/365 # 10 infected immigrant per 100,000 individuals per year
  output <- mseirv_static_sim(this_state=this_state, tf=tf, sim_type=1, sim_num=sim_num, 
                              routine_cov=routine_cov/100, ext_immi_rate=ext_immi_rate, simplified_output=TRUE)
  file_name <- paste(this_state, paste0("tf", tf), paste0("cov", routine_cov), paste0("simNum", sim_num), sep="_")
  saveRDS(output, file=file.path(LANDSCAPE_SIMS_FOLDER, file_name))
}
```

At the state level...

```{r eval=FALSE}
routine_covs <- seq(0, 100, 1) 
input <- data.frame(state=names(state_median_r0_idxs), sim_num=state_median_r0_idxs)
input <- merge(data.frame(routine_cov=routine_covs), input, by=NULL)
(parallel::mcmapply(landscape_simulator, as.character(input[["state"]]), input[["routine_cov"]], input[["sim_num"]], mc.cores=5))
```

At the national level...

```{r eval=FALSE}
routine_covs <- seq(0, 100, 1) 
input <- data.frame(state="Nigeria", sim_num=median_r0_idx)
input <- merge(data.frame(routine_cov=routine_covs), input, by=NULL)
(parallel::mcmapply(landscape_simulator, as.character(input[["state"]]), input[["routine_cov"]], input[["sim_num"]], mc.cores=1))
```

## Dynamic risk assessment simulations

Next, we run simulations where the routine vaccination coverage is allowed to change over time and where supplemental immunization campaigns are conducted among children after introduction. 

### Improvement in routine vaccination coverage 

We start by making a simple adjustment to the previous simulations. We introduce routine vaccination at the current MCV1 coverage level and then allow the RCV coverage to increase over time. In particular, each year the proportion missed during routine vaccination is decreased by some percentage. We define improvement in this way to reflect the fact that absolute increases in routine coverage become more difficult as routine RCV coverage increases. 

We focus on four scenarios: $1\%$, $2.5\%$, $5\%$ and $10\%$. Let's run simulations for these scenarios. We let `sim_type=2` to reflect the fact that routine RCV coverage improves over time. 

```{r}
basic_dynamic_simulator <- function(this_state, sim_num, rate) {
  tf <- 30
  ext_immi_rate <- 10/10^5/365 # 10 infected immigrant per 100,000 individuals per year
  routine_cov <- as.numeric(ng1_mcv1_data[which(ng1_mcv1_data$state==this_state), "MCV1"])
  routine_covs <- rep(NA, tf)
  for (ii in seq_along(routine_covs)) {
    if (ii==1) routine_covs[ii] <- routine_cov
    else routine_covs[ii] <- (100-routine_covs[ii-1])*rate+routine_covs[ii-1]
  }
  output <- mseirv_static_sim(this_state=this_state, tf=tf, sim_type=2, sim_num=sim_num, ext_immi_rate=ext_immi_rate, 
                              routine_cov=routine_covs/100, simplified_output=TRUE)
  file_name <- paste(this_state, paste0("tf", tf), paste0("vacRate", rate*1000), paste0("simNum", sim_num), sep="_")
  saveRDS(output, file=file.path(DYNAMIC_RISK_SIMS_FOLDER, file_name))
}
```

Let's run the simulations...
```{r eval=FALSE}
states <- c(unique(pull(ng2_pop, state)), "Nigeria")
sim_nums <- 1:100
rates <- c(0.01, 0.025, 0.05, 0.1)
input <- expand.grid(state=states, sim_num=sim_nums, rate=rates)
(parallel::mcmapply(basic_dynamic_simulator, as.character(input[["state"]]), input[["sim_num"]], input[["rate"]], mc.cores=6))
```

Let's run some additional simulations for each state at the median $R_0$. 

```{r eval=FALSE}
rates <- c(0, 0.01, 0.025, 0.05, 0.1)
input <- data.frame(state=rep(c(names(state_median_r0_idxs), "Nigeria"), each=length(rates)),
                    sim_num=rep(c(state_median_r0_idxs, median_r0_idx), each=length(rates)), 
                    rate=rep(rates, length(state_median_r0_idxs)+1))
(parallel::mcmapply(basic_dynamic_simulator, as.character(input[["state"]]), input[["sim_num"]], input[["rate"]], mc.cores=2))
```

### Improvement in routine vaccination coverage with SIAs

Next, we add two additional layers to the vaccination program: (1) a starting ('catch-up') campaign in the first year of vaccination targeting children aged 1 year to 14 years and (2) the same starting campaign along with regular supplemental immunization activities (`follow-up campaigns`) targeting all children aged between 1 year and 4 years at 4-year intervals. Since we now include SIAs, there are additional steps that the simulator must take. `sim_type=3` refers to the full dynamic simulations with changes in routine vaccination, SIAs, and imported infections. 

`mseirv_dynamic_sim` is similar to `mseirv_static_sim` except that it includes a program of supplemental immunisation campaigns defined by `sia_data`. It also only permits `sim_type=3`. 

```{r}
mseirv_dynamic_sim  <- function(this_state, tf, sim_num, routine_cov, sia_data, ext_immi_rate, simplified_output=FALSE) {
  # get parameters
  basic_parameters <- basic_parameters_list[[this_state]]
  beta_matrix <- beta_matrices_list[[this_state]][[sim_num]]
  omega <- omegas_list[[this_state]][sim_num]
  
  omegas <- rep(omega, basic_parameters$num_stages)
  names(omegas) <- paste0("omega", 1:basic_parameters$num_stages, sep="")
  basic_parameters$beta <- beta_matrix
  basic_parameters$omega <- omegas
  parameters <- unlist(basic_parameters)
  names(parameters) <- gsub("^.*\\.","", names(parameters))
  parameters <- c(parameters, sim_type=3)
  
  # add routine vaccination to parameters
  pv <- routine_cov; names(pv) <- paste("pv", 1:tf, sep=""); parameters <- c(parameters, pv) 
  
  # add infected immigration rate
  parameters <- c(parameters, eps=ext_immi_rate)
  
  # get initial conditions
  initial_conditions <- initial_conditions[[this_state]]
  initial_conditions <- reorder_initial_conditions(initial_conditions, 
                                                   num_stages=parameters["num_stages"])
  classes <- c("M", "S", "E", "I", "R", "V", "1SP", "2SP", "3SP", "EP")
  initial_pop <- sapply(1:(length(initial_conditions)/length(classes)), function(x) sum(initial_conditions[paste0(classes, x)]))
  names(initial_pop) <- paste0("N", 1:length(initial_pop))
  parameters <- c(parameters, initial_pop)
  
  # organize SIA data
  initial_age_distribution <- initial_age_distributions[[this_state]]
  agegrps <- initial_age_distribution$agegrp
  sia_data %<>% 
    mutate(lAgeIdx=sapply(lAge, function(x) which(agegrps==x)), 
           uAgeIdx=sapply(uAge, function(x) which(agegrps==x)-1))
  sia_start_years <- pull(sia_data, year)

  # no SIAs
  if (length(sia_start_years)==0) {
    output <- as.data.frame(deSolve::ode(y=initial_conditions, 
                                         times=seq(0, tf*365, 1),
                                         func=get_ode_rates, parms=parameters, maxsteps=10^4))
    if (simplified_output) {
      output <- output[, c("time", paste("EP", 1:parameters["num_stages"], sep=""))]
    }
    return(output)
  }
  
  # has SIAs
  for (ii in seq_along(sia_start_years)) {
    if (ii==1 && sia_start_years[ii]!=1) {
      output <- as.data.frame(deSolve::ode(y=initial_conditions, 
                                           times=seq(0, (sia_start_years[ii]-1)*365-1, 1),
                                           func=get_ode_rates, parms=parameters, maxsteps=10^4))
      initial_conditions <- output[, -1] %>% tail(1)
      state_names <- names(initial_conditions)
      initial_conditions <- as.numeric(initial_conditions)
      names(initial_conditions) <- state_names
    }
    
    initial_conditions <- conduct_sia(y=initial_conditions,
                                      st_agegrp_idx=as.numeric(sia_data[ii, "lAgeIdx"]),
                                      end_agegrp_idx=as.numeric(sia_data[ii, "uAgeIdx"]),
                                      cov=as.numeric(sia_data[ii, "cov"]/100))
    start_time <- (sia_start_years[ii]-1)*365
    if (ii==length(sia_start_years)) end_time <- tf*365-1
    else end_time <- (sia_start_years[ii+1]-1)*365-1
    section_output <- as.data.frame(deSolve::ode(y=initial_conditions, 
                                                 times=seq(start_time, end_time, 1), 
                                                 func=get_ode_rates, parms=parameters, maxsteps=10^4))
    if (ii==1 && sia_start_years[ii]==1) output <- section_output
    else output <- rbind(output, section_output)
    
    initial_conditions <- output[, -1] %>% tail(1)
    state_names <- names(initial_conditions)
    initial_conditions <- as.numeric(initial_conditions)
    names(initial_conditions) <- state_names
  }
  if (simplified_output) {
    output <- output[, c("time", paste("EP", 1:parameters["num_stages"], sep=""))]
  }
  return(output)
}
```

Next, let's run the simulations. `advanced_dynamic_simulator` takes as arguments a state name `this_state`, a simulation number `sim_num`, a starting routine coverage (optional) `custom_routine_cov`, a rate of improvement `rate`, a program of SIAs `sia_type` and logical flags for the handling of the simulation output.  

```{r}
advanced_dynamic_simulator <- function(this_state, sim_num, custom_routine_cov=NULL, rate, sia_type, return_output=FALSE, save_output=TRUE, overwrite=FALSE) {
  tf <- 30
  ext_immi_rate <- 10/10^5/365 # 10 infected immigrant per 100,000 individuals per year
  
  # routine vaccination
  if (is.null(custom_routine_cov)) {
    routine_cov <- as.numeric(ng1_mcv1_data[which(ng1_mcv1_data$state==this_state), "MCV1"])
  } else {
    routine_cov <- custom_routine_cov
  }
  routine_covs <- rep(NA, tf)
  for (ii in seq_along(routine_covs)) {
    if (ii==1) routine_covs[ii] <- routine_cov
    else routine_covs[ii] <- (100-routine_covs[ii-1])*rate+routine_covs[ii-1]
  }
  
  # supplemental immunization activities (SIAs)
  general_sia_scenario <- general_sia_scenarios[[sia_type]]
  rcv_sias <- general_sia_scenario %>% 
    select(Year, SIA, Coverage, `Lower Age`, `Upper Age`) %>% 
    filter(SIA=="Yes") %>% 
    select(-SIA) %>% 
    transmute(year=as.numeric(Year), cov=as.numeric(Coverage), 
              lAge=as.numeric(`Lower Age`), uAge=as.numeric(`Upper Age`))
  
  if (!overwrite) {
    file_name <- paste(this_state, paste0("tf", tf), sia_type, paste0("vacRate", rate*1000), paste0("simNum", sim_num), sep="_")
    file <- file.path(DYNAMIC_RISK_SIMS_FOLDER, file_name)
    if (file.exists(file)) return()
  }
  output <- mseirv_dynamic_sim(this_state=this_state, tf=tf, sim_num=sim_num,
                               routine_cov=routine_covs/100, sia_data=rcv_sias, ext_immi_rate=ext_immi_rate, 
                               simplified_output=TRUE)
  if (save_output) {
    file_name <- paste(this_state, paste0("tf", tf), sia_type, paste0("vacRate", rate*1000), paste0("simNum", sim_num), sep="_")
    saveRDS(output, file=file.path(DYNAMIC_RISK_SIMS_FOLDER, file_name))
  }
  if (return_output) return(output)
}
```

```{r eval=FALSE}
states <- c(unique(pull(ng2_pop, state)), "Nigeria")
sim_nums <- 1:100
rates <- c(0, 0.01, 0.025, 0.05, 0.1)
sia_types <- c("sia1", "sia2")
input <- expand.grid(state=states, sim_num=sim_nums, rate=rates, sia_type=sia_types)
(parallel::mcmapply(advanced_dynamic_simulator, as.character(input[["state"]]), input[["sim_num"]], input[["rate"]], input[["sia_type"]], mc.cores=5))
```

Let's run some additional simulations for each state at the median $R_0$. 

```{r eval=FALSE}
rates <- c(0, 0.01, 0.025, 0.05, 0.1)
sia_types <- c("sia1", "sia2")
input <- data.frame(state=rep(c(names(state_median_r0_idxs), "Nigeria"), each=length(rates)),
                    sim_num=rep(c(state_median_r0_idxs, median_r0_idx), each=length(rates)), 
                    rate=rep(rates, length(state_median_r0_idxs)+1), 
                    sia_type=rep(sia_types, each=(length(state_median_r0_idxs)+1)*length(rates)))
(parallel::mcmapply(advanced_dynamic_simulator, as.character(input[["state"]]), input[["sim_num"]], input[["rate"]], input[["sia_type"]], mc.cores=5))
```